### 2023-02-04 manuscript/manuscript.qmd
1st commit

---
title: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
subtitle: A Bayesian Perspective
author:
  - name: James M Brophy
    email: james.brophy@mcgill.ca
    affiliations: 
        - id: McGill University 
          name: McGill University Health Center
          department: Centre for Health Outcomes Research (CORE) 
          address: 5252 Boul. de Maisonneuve West Room 2B.37
          city: Montreal
          state: QC
          postal-code: H4A 3S5
    attributes:
        corresponding: true
    note: JMB is a research scholar supported by Les Fonds de Recherche Québec Santé
abstract: |
  A recent randomized clinical trial reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR had similar effects on survival with a favorable neurologic outcome. Herein, it is examined whether a Bayesian perspective allows any additional insights into the interpretation of this trial.
keywords: 
  - extracorporeal CPR
  - Bayesian statistics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: To be determined
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

# Introduction

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(CPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support) can improve survival and
diminish anoxic brain injury is a current topic of research. The largest
randomized clinical trial (RCT) examining this question recently
published their results[@CPR2023a]. For the primary outcome, 30 day
survival without significant neurological deficit, the authors observed
an odds ratio of 1.4 (95% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR for leading to their conclusion "In patients
with refractory out-of-hospital cardiac arrest, extracorporeal CPR and
conventional CPR had similar effects on survival with a favorable
neurological outcome".[@CPR2023a]

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals[@RN5420]. Rather it assumes the reader has perhaps heard that
Bayesian methods mirror our intuitive learning process and is curious
about its potential application to RCT interpretations.

Therefore the goal of this communication is to examine whether a
Bayesian perspective permits additional insights into the specific
clinical question regarding any added value of extracorporeal CPR
following an out-of-hospital arrest in patients refractory to standard
CPR.

# Methods

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial [@CPR2023a] and used for
the primary analysis. The ITT analysis has the advantage of minimizing
bias by preserving the prognostic balance afforded by randomization as
well as assuring the validity of the statistical analyses. ITT assesses
subjects based on the group they were initially (and randomly) allocated
to, regardless of whether or not they dropped out, were fully adhered to
the treatment or switched to an alternative treatment. ITT analyses can
therefore be seen as a conservative estimate which mirrors clinical
effectiveness. In contrast, a per protocol (PP) analysis involves a
comparison of treatment groups in a trial that includes only those
patients who completed the treatment they were originally allocated to.
Similarly an "as treated" analysis considers only which treatments
subjects received, regardless of their randomization status and protocol
adherence. While both PP and as-treated analyses alone may lead to bias,
in conjunction with an ITT analysis they may provide additional insights
into efficacy and have also been examined from a Bayesian perspective.

Bayesian analytical approaches provide a number of benefits over the
classical NHST approach, including parameter estimation accompanied by
direct probability statements about parameters of interest (herein the
risk of survival with intact neurological status), and the incorporation
prior knowledge [@BrophyCardio; @Zampieri].

These probability statements arise from the posterior distribution
according to the following equation:
$$ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} $$

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function) one requires a prior probability
distribution for each parameter. The mechanics of the Bayesian analyses
were performed using the Stan programming language [@stan] through the R
package rstanarm [@rstanarm] and fit a logistic regression model with a
single treatment parameter, $\theta$. Because our focus is the
interpretation of the INCEPTION trial alone, our primary analysis used
rstanarm's default vague parameter priors
($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
posterior distribution is dominated by the observed INCEPTION data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Using prior
information also underscores the important advantage of Bayesian
analyses to learn sequentially. There were two previous RCTs examining
extracorporeal CPR[@RN6759][@RN6751] and while the protocols are not
identical, it may be reasonable to allow this data can serve as an
informed prior which can be updated with the INCEPTION data. This prior
information of the probability of success in each arm, $X_i$, can be summarized by a Beta distribution as follow
$$X_i\sim\mbox{Beta}(\alpha_0,\beta_0)$$ 
where in the context of Bernoulli trials, $\alpha$ can be interpreted as 1 + number of successes and $\beta$ can be interpreted as 1 + number of failures. The mean and variance of this distribution are 
$$\mu=\frac{\alpha}{\alpha+\beta}$$

$$\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}$$ where
$\alpha$ and $\beta$ are the number of successes and failures respectively. The difference between the two treatment arms of these previous studies can then serve as the prior information which in the Bayesian paradigm can be updated with the INCEPTION data to form a posterior distribution.

Posterior distributions are summarized with medians and 95%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95% of the probability density function
[@mcelreath2020]. We not only calculated the posterior probability of
any benefit (OR \>1.00), but also of clinically meaningful benefits
(defined as OR \>1.10).

All analyses were executed within the integrated development environment
of RStudio and the statistical code can be found on Github
(https://github.com/brophyj/eCPR).

# Results

The ITT data from the three pertinent trials is shown in Table 1. Performing a Bayesian analysis on the INCEPTION trial, using a vague prior, produces an odds ratio 1.32 (95% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result to the original analysis confirms the lack if impact of the vague prior and shows this Bayesian analysis is dominated by the observed data.

\newpage

```{r echo=FALSE,  message=FALSE, warning=FALSE, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 1, digits = 3)
pander::panderOptions('keep.trailing.zeros', TRUE)

library(here)
library(tidyverse)
library(broom)
library(cobalt)
library(table1)
library(knitr)
library(patchwork)
library(rstanarm)
library(tidybayes)

```

# Tables

```{r}
df <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
             fail = c(52, 56, 15, 8, 108, 86),
             success = c(10,14, 0, 6,24,38 )) %>% 
  mutate(total = fail + success,
         prop_success = success / total) 
df1 <- df[, c(1:4)]  %>% tidyr::pivot_wider(names_from = Tx, values_from = c(fail,success))

knitr::kable(df1, caption = "Table 1 Extracted ITT trial data", col.names=c("Trial", "Fail CPR", "Fail eCPR", "Success CPR", "Success eCPR"))

#frequentist INCEPTION  to reproduce NEJM result
fit <- glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total)
# summary(fit, digits=3)
# print(c(exp(.262),exp(.262+c(-1,1)*1.96*.457)))
# [1] 1.300 0.531 3.183
```

# Bayesian analysis

```{r cache=TRUE}
# - INCEPTION alone default vague priors
set.seed(1234)
fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0)
# summary(fit1, digits=3)
# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
# get_variables(fit1)
# prior_summary(fit1)
# help('prior_summary.stanreg')

# informative priors using 2 previous studies
# one method
t1 <- rbeta(100000,25,124)
t2 <- rbeta(100000,45,95)
t3 <- t2 - t1
# t3 can be approximated by a normal distribution with the following parameters
# mean(t3)
# [1] 0.154
# sd(t3)
# [1] 0.0498

# second method
library(LearnBayes)
quantile1=list(p=.025, x=quantile(t3,0.025))     # 2.5% quantile should be 0.0556 
quantile2=list(p=.975, x=quantile(t3,0.975))      # 97.5% quantile should be 0..251
par_beta <- beta.select(quantile1, quantile2)
m_beta <-  par_beta[1]/(par_beta[1]+par_beta[2])
v_beta <- (par_beta[1] * par_beta[2])/((par_beta[1]+par_beta[2])^2 * (par_beta[1] +par_beta[2] + 1))
sd_beta <- sqrt(v_beta)
# Beta (6.35, 39.54) gives mean = 0.138 sd = 0.0504 close to estimates using method 1

# stan_glm doesn't seem able to take a Beta prior distribution so will use normal approximation

fit2 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0,
                 prior = normal(mean(t3), sd(t3)))
summary(fit2, digits=4)

```

**Have obviously not done this correctly as the prior should have shifted the default model parameter for eCPR to a larger number with a small sd**

# Figures

```{r}

# tt <- fit1 %>%
#  spread_draws(`(Intercept)`, `TxeCPR`)
# sum(tt$TxeCPR>0)/nrow(tt)

fit1 %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior") +
  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 71%") +
  theme_classic() +
  theme(legend.position="none")

# Examining prior information
x <- seq(0, .5, by = 0.001)

#THEN EACH OF OUR BETAS DISTR

beta1 <- dbeta(x, shape1 = 25, shape2 = 124) # probability of success with CPR
beta2 <- dbeta(x, shape1 = 45, shape2 = 95)  # probability of success with eCPR
beta3 <- dbeta(x, shape1 = 6.36, shape2 = 39.65)

mixture <- ggplot()+
  geom_line(aes(x, beta1),  color = "red",
            linewidth = 1.2) +
  geom_line(aes(x, beta2), color = "green",
            linewidth = 1.2) +
  geom_line(aes(x, beta3), color = "black",
            linewidth = 1.2) +
  labs(title = "Figure 2 Beta Distribution of prior studies",
       x = "Sequence",
       y = "Beta",
       caption = "Green = p(+outcome | eCPR) \nRed = p(+outcome | CPR) \nBlack= difference of the 2 probabilities")

mixture

```

\newpage

# References {.unnumbered}

### 2023-02-04 manuscript/manuscript.tex
1st commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
  pdfauthor={James M Brophy},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac
Arrest \\\large{A Bayesian Perspective} }
\author[1]{James M Brophy%
\corref{cor1}%
\fnref{fn1}}
 \ead{james.brophy@mcgill.ca}

\affiliation[1]{organization={McGill University Health Center, Centre
for Health Outcomes Research (CORE)},addressline={5252 Boul. de
Maisonneuve West Room 2B.37},city={Montreal},postcode={H4A
3S5},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{JMB is a research scholar supported by Les Fonds de
Recherche Québec Santé}

\begin{abstract}
A recent randomized clinical trial reported in patients with refractory
out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR
had similar effects on survival with a favorable neurologic outcome.
Herein, it is examined whether a Bayesian perspective allows any
additional insights into the interpretation of this trial.
\end{abstract}

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(CPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support) can improve survival and
diminish anoxic brain injury is a current topic of research. The largest
randomized clinical trial (RCT) examining this question recently
published their results\citep{CPR2023a}. For the primary outcome, 30 day
survival without significant neurological deficit, the authors observed
an odds ratio of 1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR for leading to their conclusion ``In
patients with refractory out-of-hospital cardiac arrest, extracorporeal
CPR and conventional CPR had similar effects on survival with a
favorable neurological outcome''.\citep{CPR2023a}

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals\citep{RN5420}. Rather it assumes the reader has perhaps heard
that Bayesian methods mirror our intuitive learning process and is
curious about its potential application to RCT interpretations.

Therefore the goal of this communication is to examine whether a
Bayesian perspective permits additional insights into the specific
clinical question regarding any added value of extracorporeal CPR
following an out-of-hospital arrest in patients refractory to standard
CPR.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial \citep{CPR2023a} and used
for the primary analysis. The ITT analysis has the advantage of
minimizing bias by preserving the prognostic balance afforded by
randomization as well as assuring the validity of the statistical
analyses. ITT assesses subjects based on the group they were initially
(and randomly) allocated to, regardless of whether or not they dropped
out, were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis involves a comparison of treatment groups in a trial that
includes only those patients who completed the treatment they were
originally allocated to. Similarly an ``as treated'' analysis considers
only which treatments subjects received, regardless of their
randomization status and protocol adherence. While both PP and
as-treated analyses alone may lead to bias, in conjunction with an ITT
analysis they may provide additional insights into efficacy and have
also been examined from a Bayesian perspective.

Bayesian analytical approaches provide a number of benefits over the
classical NHST approach, including parameter estimation accompanied by
direct probability statements about parameters of interest (herein the
risk of survival with intact neurological status), and the incorporation
prior knowledge \citep{BrophyCardio, Zampieri}.

These probability statements arise from the posterior distribution
according to the following equation:
\[ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} \]

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function) one requires a prior probability
distribution for each parameter. The mechanics of the Bayesian analyses
were performed using the Stan programming language \citep{stan} through
the R package rstanarm \citep{rstanarm} and fit a logistic regression
model with a single treatment parameter, \(\theta\). Because our focus
is the interpretation of the INCEPTION trial alone, our primary analysis
used rstanarm's default vague parameter priors
(\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring that the
posterior distribution is dominated by the observed INCEPTION data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Using prior
information also underscores the important advantage of Bayesian
analyses to learn sequentially. There were two previous RCTs examining
extracorporeal CPR{[}\citet{RN6759}{]}\citep{RN6751} and while the
protocols are not identical, it may be reasonable to allow this data can
serve as an informed prior which can be updated with the INCEPTION data.
This prior information of the probability of success in each arm,
\(X_i\), can be summarized by a Beta distribution as follow
\[X_i\sim\mbox{Beta}(\alpha_0,\beta_0)\] where in the context of
Bernoulli trials, \(\alpha\) can be interpreted as 1 + number of
successes and \(\beta\) can be interpreted as 1 + number of failures.
The mean and variance of this distribution are
\[\mu=\frac{\alpha}{\alpha+\beta}\]

\[\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}\] where
\(\alpha\) and \(\beta\) are the number of successes and failures
respectively. The difference between the two treatment arms of these
previous studies can then serve as the prior information which in the
Bayesian paradigm can be updated with the INCEPTION data to form a
posterior distribution.

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. We not only calculated the posterior probability
of any benefit (OR \textgreater1.00), but also of clinically meaningful
benefits (defined as OR \textgreater1.10).

All analyses were executed within the integrated development environment
of RStudio and the statistical code can be found on Github
(https://github.com/brophyj/eCPR).

\hypertarget{results}{%
\section{Results}\label{results}}

The ITT data from the three pertinent trials is shown in Table 1.
Performing a Bayesian analysis on the INCEPTION trial, using a vague
prior, produces an odds ratio 1.32 (95\% Credible Interval (CrI) 0.54 -
3.22). The closeness of this result to the original analysis confirms
the lack if impact of the vague prior and shows this Bayesian analysis
is dominated by the observed data.

\newpage

\hypertarget{tables}{%
\section{Tables}\label{tables}}

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Table 1 Extracted ITT trial data}\tabularnewline
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endfirsthead
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endhead
INCEPTION & 52 & 56 & 10 & 14 \\
ARREST & 15 & 8 & 0 & 6 \\
PRAGUE & 108 & 86 & 24 & 38 \\
\bottomrule()
\end{longtable}

\hypertarget{bayesian-analysis}{%
\section{Bayesian analysis}\label{bayesian-analysis}}

\begin{verbatim}

Attaching package: 'LearnBayes'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:rstanarm':

    laplace
\end{verbatim}

\begin{verbatim}

Model Info:
 function:     stan_glm
 family:       binomial [logit]
 formula:      prop_success ~ Tx
 algorithm:    sampling
 sample:       4000 (posterior sample size)
 priors:       see help('prior_summary')
 observations: 2
 predictors:   2

Estimates:
              mean    sd      10%     50%     90%  
(Intercept) -1.5987  0.2265 -1.8906 -1.5963 -1.3088
TxeCPR       0.1542  0.0501  0.0897  0.1535  0.2181

Fit Diagnostics:
           mean    sd      10%     50%     90%  
mean_PPD 11.9740  3.1650  8.0000 12.0000 16.0000

The mean_ppd is the sample average posterior predictive distribution of the outcome variable (for details see help('summary.stanreg')).

MCMC diagnostics
              mcse   Rhat   n_eff
(Intercept)   0.0042 0.9995 2872 
TxeCPR        0.0009 0.9999 2827 
mean_PPD      0.0544 0.9994 3387 
log-posterior 0.0238 1.0006 1835

For each parameter, mcse is Monte Carlo standard error, n_eff is a crude measure of effective sample size, and Rhat is the potential scale reduction factor on split chains (at convergence Rhat=1).
\end{verbatim}

\textbf{Have obviously not done this correctly as the prior should have
shifted the default model parameter for eCPR to a larger number with a
small sd}

\hypertarget{figures}{%
\section{Figures}\label{figures}}

\begin{verbatim}
Warning: Removed 9 rows containing missing values (`stat_slabinterval()`).
\end{verbatim}

\begin{verbatim}
Warning: Using the `size` aesthietic with geom_segment was deprecated in ggplot2 3.4.0.
i Please use the `linewidth` aesthetic instead.
\end{verbatim}

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-2.pdf}

\newpage

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2023-02-04 analysis.R

@@ -1,0 +1,140 @@
+cbind(switch, n - switch) ~ dist100 + arsenic
+
+df <- data.frame(dead=c(56, 52), n=c(70,62), Rx=c(1,0))
+df
+glm(cbind(dead, n-dead) ~ Rx, data=df, family = binomial, weights = n )
+
+
+library(dplyr)
+df <- tibble(Tx = c("CPR", "eCPR"),
+             dead = c(52, 56),
+             alive = c(10,14)) %>% 
+  mutate(total = dead + alive,
+         prop_alive = alive / total) 
+
+fit <- glm(prop_alive ~ Tx, data = df, family = binomial(link="logit"), weights = total)
+summary(fit, digits=3)
+print(c(exp(.262),exp(.262+c(-1,1)*1.96*.457)))
+
+df
+
+
+library(rstanarm)
+fit1 <- stan_glm(prop_alive ~ Tx, data = df, family = binomial(link="logit"), weights = total, refresh=0)
+summary(fit1, digits=4)
+
+library(tidybayes)
+get_variables(fit1)
+prior_summary(fit1)
+help('prior_summary.stanreg') 
+
+fit1 %>%
+  spread_draws(`(Intercept)`, `TxeCPR`) %>%
+  head(10)
+
+posterior_vs_prior(fit1, pars = "beta")
+plot(fit1)
+
+fit1 %>% 
+  tidybayes::tidy_draws() %>% 
+  ggdist::mean_qi(exp(TxeCPR))
+
+fit1 %>%
+  spread_draws(`(Intercept)`, `TxeCPR`) %>%
+  median_qi(`(Intercept)`, `TxeCPR`)
+
+
+fit1 %>%
+  spread_draws(`(Intercept)`, `TxeCPR`) %>%
+  median_qi(condition_mean = `(Intercept)` + `TxeCPR`, .width = c(.95, .68)) %>%
+  ggplot(aes(y = `TxeCPR`, x = condition_mean, xmin = .lower, xmax = .upper)) +
+  geom_pointinterval()
+
+fit1 %>% tidybayes::tidy_draws() |> 
+  ggplot(aes(x = exp(TxeCPR),
+             fill = after_stat(x > 1),
+             slab_alpha = after_stat((x)))
+  ) +
+  ggdist::stat_halfeye() +
+  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_y_continuous(name="Density") +
[…]

### 2023-02-07 manuscript/manuscript.qmd
updated priors

@@
 set.seed(1234)
-fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0)
+fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
+fit1_sum <- summary(fit1, digits=3)
+conf_95 <- c(exp(fit1_sum[2]),exp(fit1_sum[2]+c(-1,1)*1.96*fit1_sum[10]))
+tt <- fit1 %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
 # summary(fit1, digits=3)
@@
 
-# informative priors using 2 previous studies
-# one method
-t1 <- rbeta(100000,25,124)
-t2 <- rbeta(100000,45,95)
-t3 <- t2 - t1
-# t3 can be approximated by a normal distribution with the following parameters
-# mean(t3)
-# [1] 0.154
-# sd(t3)
-# [1] 0.0498
+#################
+# function to calculate point estimate, 95% CrI, and P(HR>1)
+# using prior information
+################
+eCRP_bayes <- function(data, m_CPR, sd_CPR, m_eCPR, sd_eCPR ) {
+  set.seed(1234)
+  fit <- stan_glm(prop_success ~ Tx, data = data, family = binomial(link="logit"), weights = total, refresh=0,
+                  prior = normal(m_eCPR, sd_eCPR),
+                  prior_intercept = normal (m_CPR, sd_CPR))
+  fit_sum <- summary(fit, digits=3)
+  conf_95 <- c(exp(fit_sum[2]),exp(fit_sum[2]+c(-1,1)*1.96*fit_sum[10]))
+  tt <- fit %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+  fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
+  fit_draws <- fit %>% tidybayes::tidy_draws() 
+  my_list <- list(fig_prob, conf_95, fit_draws)
+  return(my_list)
+}
 
-# second method
-library(LearnBayes)
-quantile1=list(p=.025, x=quantile(t3,0.025))     # 2.5% quantile should be 0.0556 
-quantile2=list(p=.975, x=quantile(t3,0.975))      # 97.5% quantile should be 0..251
-par_beta <- beta.select(quantile1, quantile2)
-m_beta <-  par_beta[1]/(par_beta[1]+par_beta[2])
-v_beta <- (par_beta[1] * par_beta[2])/((par_beta[1]+par_beta[2])^2 * (par_beta[1] +par_beta[2] + 1))
-sd_beta <- sqrt(v_beta)
-# Beta (6.35, 39.54) gives mean = 0.138 sd = 0.0504 close to estimates using method 1
+# combined_prior 
+# informative priors using 2 previous studies combined
+comb_m_CPR <- sum(df1[c(2,3),4]) / (sum(df1[c(2,3),4]) + sum(df1[c(2,3),2])) # proportion previous results for CPR
+comb_sd_CPR <- sqrt(comb_m_CPR*(1-comb_m_CPR))                               
+comb_m_eCPR <- sum(df1[c(2,3),5]) / (sum(df1[c(2,3),5]) + sum(df1[c(2,3),3])) # proportion previous results for eCPR
+comb_sd_eCPR <- sqrt(comb_m_eCPR*(1-comb_m_eCPR))
+comb_prior <- eCRP_bayes(df[c(1:2),], comb_m_CPR, comb_sd_CPR, comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
 
-# stan_glm doesn't seem able to take a Beta prior distribution so will use normal approximation
+# enthusiastic prior with only ARREST study
+enth_m_CPR <- sum(df1[c(2),4]) + 1 / (sum(df1[c(2),4]) + sum(df1[c(2),2])) # proportion previous results for CPR, + 0.5 to avoid 0 cell
+enth_sd_CPR <- sqrt(enth_m_CPR*(1-enth_m_CPR))
+enth_m_eCPR <- sum(df1[c(2),5]) / (sum(df1[c(2),5]) + sum(df1[c(2),3])) # proportion previous results for eCPR
+enth_sd_eCPR <- sqrt(enth_m_eCPR*(1-enth_m_eCPR))
+enth_prior <- eCRP_bayes(df[c(1:2),], enth_m_CPR, enth_sd_CPR, enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
 
-fit2 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0,
-                 prior = normal(mean(t3), sd(t3)))
-summary(fit2, digits=4)
+# skeptical prior with only PRAGUE study
+skep_m_CPR <- sum(df1[c(3),4])  / (sum(df1[c(3),4]) + sum(df1[c(3),2])) # proportion previous results for CPR
+skep_sd_CPR <- sqrt(skep_m_CPR*(1-skep_m_CPR))
+skep_m_eCPR <- sum(df1[c(3),5]) / (sum(df1[c(3),5]) + sum(df1[c(3),3])) # proportion previous results for eCPR
+skep_sd_eCPR <- sqrt(skep_m_eCPR*(1-skep_m_eCPR))
+skep_prior <- eCRP_bayes(df[c(1:2),], skep_m_CPR, skep_sd_CPR, skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior
 
+# skep_prior <- eCRP_bayes(df[c(1:2),], enth_m_CPR, enth_sd_CPR, skep_m_eCPR, skep_sd_eCPR) #test to see impact of enth intercept
+
+table_tbl <- 
+  tibble(
+    prior = c("vague", "combined", "enthusiastic", "skeptical"),
+    pt_est <- c(conf_95[1], comb_prior[[2]][1], enth_prior[[2]][1], skep_prior[[2]][1]),
+    llimit <- c(conf_95[2], comb_prior[[2]][2], enth_prior[[2]][2], skep_prior[[2]][2]),
+    ulimit <- c(conf_95[3], comb_prior[[2]][3], enth_prior[[2]][3], skep_prior[[2]][3]),
+    p.gt.1 <- c(fig_prob, comb_prior[1], enth_prior[1], skep_prior[1])
+  )
+
+library(gt)
+gt(table_tbl) %>%
+  tab_stubhead(label = "Priors") 
 ```
 
-**Have obviously not done this correctly as the prior should have shifted the default model parameter for eCPR to a larger number with a small sd**
+# **Summary of what I don't understand**
+
+This makes only partial sense.
+95% CrI are narrower with informative priors.
+OK, but one would think the enthusiastic prior from a trial ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes for eCPR vs CPR would shift the HR of 1.32 to the right but instead the HR falls to have a 1.27.
+The skeptical prior from PRAGUE which was stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as expected
+
+So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)    
+skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)
+
+Tx for enthusiastic prior > Tx skeptical prior (eCPR) but this seems to be offset by the reverse occurring for the intercept (CPR).     
+The problem seems to be in how comparable are these previous RCTs as the baseline success for the standard TX (CRP) is radically different 0/14 vs 24/132.
+
+When I use a prior that uses the enthusiastic intercept prior and the skeptical Tx prior, I  get a bigger downward shift with HR 1.220 95% CrI(0.678 2.194) and P>1 74.5%
+
+Can you see any other explanation?
+
+Do you have any other insights?
 
@@
 
-```{r}
+```{r warning=FALSE, message=FALSE}
 
-# tt <- fit1 %>%
-#  spread_draws(`(Intercept)`, `TxeCPR`)
-# sum(tt$TxeCPR>0)/nrow(tt)
+#############
+# function to plot probability function
+#############
+
+plot_bayes <- function(data,text1,text2, fig_prob=.5){
+  annon <- paste0(text2, fig_prob*100, "%")
+  ggplot(data, aes(x = exp(TxeCPR),
+                   fill = after_stat(x > 1),
+                   slab_alpha = after_stat((x)))
+  ) +
+    ggdist::stat_halfeye() +
+    scale_x_continuous(name=text, limits=c(0, 4)) +
+    scale_y_continuous(name="Density") +
+    labs(title = text1) +
+    annotate(geom="text", x=2.75,y=.6, label = annon) +
+    theme_classic() +
+    theme(legend.position="none") 
+}
+
+#plot combined prior
+#text1 <-c("HR increased survival with extracorporeal CPR")
+#text2 <- c("Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR >")
+#plot_bayes(comb_prior[[3]], text1, text2, comb_prior[[1]])
 
@@
   labs(title = "Figure 1 INCEPTION ITT analysis with vague prior") +
-  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 71%") +
+  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7%") +
   theme_classic() +
   theme(legend.position="none")
-
-# Examining prior information
-x <- seq(0, .5, by = 0.001)
-
-#THEN EACH OF OUR BETAS DISTR
-
-beta1 <- dbeta(x, shape1 = 25, shape2 = 124) # probability of success with CPR
-beta2 <- dbeta(x, shape1 = 45, shape2 = 95)  # probability of success with eCPR
-beta3 <- dbeta(x, shape1 = 6.36, shape2 = 39.65)
-
-mixture <- ggplot()+
-  geom_line(aes(x, beta1),  color = "red",
-            linewidth = 1.2) +
-  geom_line(aes(x, beta2), color = "green",
-            linewidth = 1.2) +
-  geom_line(aes(x, beta3), color = "black",
-            linewidth = 1.2) +
-  labs(title = "Figure 2 Beta Distribution of prior studies",
-       x = "Sequence",
-       y = "Beta",
-       caption = "Green = p(+outcome | eCPR) \nRed = p(+outcome | CPR) \nBlack= difference of the 2 probabilities")
-
-mixture
 

### 2023-02-07 manuscript/manuscript.tex
updated priors

@@
 
+\usepackage{amsmath}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
 \makeatletter
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{verbatim}
+\begin{longtable}{lrrrc}
+\toprule
+prior & pt\_est <- ... & llimit <- ... & ulimit <- ... & p.gt.1 <- ... \\ 
+\midrule
+vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
+combined & 1.27 & 0.696 & 2.31 & 0.782 \\ 
+enthusiastic & 1.27 & 0.708 & 2.27 & 0.79 \\ 
+skeptical & 1.25 & 0.695 & 2.26 & 0.777 \\ 
+\bottomrule
+\end{longtable}
 
-Attaching package: 'LearnBayes'
-\end{verbatim}
+\hypertarget{summary-of-what-i-dont-understand}{%
+\section{\texorpdfstring{\textbf{Summary of what I don't
+understand}}{Summary of what I don't understand}}\label{summary-of-what-i-dont-understand}}
 
-\begin{verbatim}
-The following object is masked from 'package:rstanarm':
+This makes only partial sense. 95\% CrI are narrower with informative
+priors. OK, but one would think the enthusiastic prior from a trial
+ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes
+for eCPR vs CPR would shift the HR of 1.32 to the right but instead the
+HR falls to have a 1.27. The skeptical prior from PRAGUE which was
+stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as
+expected
 
-    laplace
-\end{verbatim}
+So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)\\
+skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)
 
-\begin{verbatim}
+Tx for enthusiastic prior \textgreater{} Tx skeptical prior (eCPR) but
+this seems to be offset by the reverse occurring for the intercept
+(CPR).\\
+The problem seems to be in how comparable are these previous RCTs as the
+baseline success for the standard TX (CRP) is radically different 0/14
+vs 24/132.
 
-Model Info:
- function:     stan_glm
- family:       binomial [logit]
- formula:      prop_success ~ Tx
- algorithm:    sampling
- sample:       4000 (posterior sample size)
- priors:       see help('prior_summary')
- observations: 2
- predictors:   2
+When I use a prior that uses the enthusiastic intercept prior and the
+skeptical Tx prior, I get a bigger downward shift with HR 1.220 95\%
+CrI(0.678 2.194) and P\textgreater1 74.5\%
 
-Estimates:
-              mean    sd      10%     50%     90%  
-(Intercept) -1.5987  0.2265 -1.8906 -1.5963 -1.3088
-TxeCPR       0.1542  0.0501  0.0897  0.1535  0.2181
+Can you see any other explanation?
 
-Fit Diagnostics:
-           mean    sd      10%     50%     90%  
-mean_PPD 11.9740  3.1650  8.0000 12.0000 16.0000
-
-The mean_ppd is the sample average posterior predictive distribution of the outcome variable (for details see help('summary.stanreg')).
-
-MCMC diagnostics
-              mcse   Rhat   n_eff
-(Intercept)   0.0042 0.9995 2872 
-TxeCPR        0.0009 0.9999 2827 
-mean_PPD      0.0544 0.9994 3387 
-log-posterior 0.0238 1.0006 1835
-
-For each parameter, mcse is Monte Carlo standard error, n_eff is a crude measure of effective sample size, and Rhat is the potential scale reduction factor on split chains (at convergence Rhat=1).
-\end{verbatim}
-
-\textbf{Have obviously not done this correctly as the prior should have
-shifted the default model parameter for eCPR to a larger number with a
-small sd}
+Do you have any other insights?
 
@@
 
-\begin{verbatim}
-Warning: Removed 9 rows containing missing values (`stat_slabinterval()`).
-\end{verbatim}
-
-\begin{verbatim}
-Warning: Using the `size` aesthietic with geom_segment was deprecated in ggplot2 3.4.0.
-i Please use the `linewidth` aesthetic instead.
-\end{verbatim}
-
 \includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}
-
-\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-2.pdf}
 

### 2023-02-07 analysis.R

@@ -1,95 +1,3 @@
-cbind(switch, n - switch) ~ dist100 + arsenic
-
-df <- data.frame(dead=c(56, 52), n=c(70,62), Rx=c(1,0))
-df
-glm(cbind(dead, n-dead) ~ Rx, data=df, family = binomial, weights = n )
-
-
-library(dplyr)
-df <- tibble(Tx = c("CPR", "eCPR"),
-             dead = c(52, 56),
-             alive = c(10,14)) %>% 
-  mutate(total = dead + alive,
-         prop_alive = alive / total) 
-
-fit <- glm(prop_alive ~ Tx, data = df, family = binomial(link="logit"), weights = total)
-summary(fit, digits=3)
-print(c(exp(.262),exp(.262+c(-1,1)*1.96*.457)))
-
-df
-
-
-library(rstanarm)
-fit1 <- stan_glm(prop_alive ~ Tx, data = df, family = binomial(link="logit"), weights = total, refresh=0)
-summary(fit1, digits=4)
-
-library(tidybayes)
-get_variables(fit1)
-prior_summary(fit1)
-help('prior_summary.stanreg') 
-
-fit1 %>%
-  spread_draws(`(Intercept)`, `TxeCPR`) %>%
-  head(10)
-
-posterior_vs_prior(fit1, pars = "beta")
-plot(fit1)
-
-fit1 %>% 
-  tidybayes::tidy_draws() %>% 
-  ggdist::mean_qi(exp(TxeCPR))
-
-fit1 %>%
-  spread_draws(`(Intercept)`, `TxeCPR`) %>%
-  median_qi(`(Intercept)`, `TxeCPR`)
-
-
-fit1 %>%
-  spread_draws(`(Intercept)`, `TxeCPR`) %>%
-  median_qi(condition_mean = `(Intercept)` + `TxeCPR`, .width = c(.95, .68)) %>%
-  ggplot(aes(y = `TxeCPR`, x = condition_mean, xmin = .lower, xmax = .upper)) +
-  geom_pointinterval()
-
-fit1 %>% tidybayes::tidy_draws() |> 
-  ggplot(aes(x = exp(TxeCPR),
-             fill = after_stat(x > 1),
-             slab_alpha = after_stat((x)))
-  ) +
-  ggdist::stat_halfeye() +
-  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
[…]

### 2023-02-07 manuscript/manuscript.qmd
2nd Bayesian with priors only for Tx

@@
 library(tidybayes)
+library(gt)
 
@@
 
-library(gt)
 gt(table_tbl) %>%
@@
 Do you have any other insights?
+
+Update 2023/02/07
+Might make most sense to interpret the INCEPTION trial based on the baseline CPR risk observed in that trial and provide informative priors only for the Tx variable (enthusiastic from ARREST, skeptical from PRAGUE)
+
+# Bayesian prior analysis #2
+```{r}
+## try re-running using prior for eCPR from different studies but leaving CPR with vague prior since baseline risks with CPR seem to vary
+## widely btw the different studies
+
+eCRP_bayes_no_intercept_prior <- function(data, m_eCPR, sd_eCPR ) {
+  set.seed(1234)
+  fit <- stan_glm(prop_success ~ Tx, data = data, family = binomial(link="logit"), weights = total, refresh=0,
+                  prior = normal(m_eCPR, sd_eCPR))
+  fit_sum <- summary(fit, digits=3)
+  conf_95 <- c(exp(fit_sum[2]),exp(fit_sum[2]+c(-1,1)*1.96*fit_sum[10]))
+  tt <- fit %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+  fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
+  fit_draws <- fit %>% tidybayes::tidy_draws() 
+  my_list <- list(fig_prob, conf_95, fit_draws)
+  return(my_list)
+}
+
+comb_prior1 <-eCRP_bayes_no_intercept_prior(df[c(1:2),], comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
+enth_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
+skep_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior
+
+table_tbl_1 <- 
+  tibble(
+    prior = c("vague", "combined", "enthusiastic", "skeptical"),
+    pt_est <- c(conf_95[1], comb_prior1[[2]][1], enth_prior1[[2]][1], skep_prior1[[2]][1]),
+    llimit <- c(conf_95[2], comb_prior1[[2]][2], enth_prior1[[2]][2], skep_prior1[[2]][2]),
+    ulimit <- c(conf_95[3], comb_prior1[[2]][3], enth_prior1[[2]][3], skep_prior1[[2]][3]),
+    p.gt.1 <- c(fig_prob, comb_prior[1], enth_prior[1], skep_prior[1])
+  )
+
+gt(table_tbl_1) %>%
+  tab_stubhead(label = "Priors") 
+```
+
+This seems like a reasonable solution
 

### 2023-02-07 manuscript/manuscript.tex
2nd Bayesian with priors only for Tx

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
+Update 2023/02/07 Might make most sense to interpret the INCEPTION trial
+based on the baseline CPR risk observed in that trial and provide
+informative priors only for the Tx variable (enthusiastic from ARREST,
+skeptical from PRAGUE)
+
+\hypertarget{bayesian-prior-analysis-2}{%
+\section{Bayesian prior analysis \#2}\label{bayesian-prior-analysis-2}}
+
+\begin{longtable}{lrrrc}
+\toprule
+prior & pt\_est <- ... & llimit <- ... & ulimit <- ... & p.gt.1 <- ... \\ 
+\midrule
+vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
+combined & 1.35 & 0.705 & 2.58 & 0.782 \\ 
+enthusiastic & 1.40 & 0.738 & 2.67 & 0.79 \\ 
+skeptical & 1.32 & 0.700 & 2.50 & 0.777 \\ 
+\bottomrule
+\end{longtable}
+
+This seems like a reasonable solution
+
 \hypertarget{figures}{%
@@
 
-\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-5-1.pdf}
 

### 2023-02-08 manuscript/manuscript.qmd
updated Tables and Graphs

--- 2023-02-04 manuscript/manuscript.qmd
@@
 
+```{r echo=FALSE,  message=FALSE, warning=FALSE, include=FALSE, results='hide'}
+knitr::opts_chunk$set(echo = FALSE)
+options(scipen = 1, digits = 3)
+pander::panderOptions('keep.trailing.zeros', TRUE)
+
+library(here)
+library(tidyverse)
+library(broom)
+library(cobalt)
+library(table1)
+library(knitr)
+library(patchwork)
+library(rstanarm)
+library(tidybayes)
+library(gt)
+
+```
+
 # Introduction
@@
 extracorporeal CPR (the addition of extracorporeal membrane oxygenation
-to standard advanced cardiac life support) can improve survival and
+to standard advanced cardiac life support (eCPR)) can improve survival and
 diminish anoxic brain injury is a current topic of research. The largest
@@
 bias by preserving the prognostic balance afforded by randomization as
-well as assuring the validity of the statistical analyses. ITT assesses
-subjects based on the group they were initially (and randomly) allocated
-to, regardless of whether or not they dropped out, were fully adhered to
-the treatment or switched to an alternative treatment. ITT analyses can
-therefore be seen as a conservative estimate which mirrors clinical
-effectiveness. In contrast, a per protocol (PP) analysis involves a
-comparison of treatment groups in a trial that includes only those
-patients who completed the treatment they were originally allocated to.
-Similarly an "as treated" analysis considers only which treatments
-subjects received, regardless of their randomization status and protocol
-adherence. While both PP and as-treated analyses alone may lead to bias,
-in conjunction with an ITT analysis they may provide additional insights
-into efficacy and have also been examined from a Bayesian perspective.
+well as assuring the validity of the statistical analyses.
 
@@
 analyses to learn sequentially. There were two previous RCTs examining
-extracorporeal CPR[@RN6759][@RN6751] and while the protocols are not
-identical, it may be reasonable to allow this data can serve as an
-informed prior which can be updated with the INCEPTION data. This prior
-information of the probability of success in each arm, $X_i$, can be summarized by a Beta distribution as follow
-$$X_i\sim\mbox{Beta}(\alpha_0,\beta_0)$$ 
-where in the context of Bernoulli trials, $\alpha$ can be interpreted as 1 + number of successes and $\beta$ can be interpreted as 1 + number of failures.
-The mean and variance of this distribution are
-$$\mu=\frac{\alpha}{\alpha+\beta}$$
+extracorporeal CPR[@RN6759; @RN6751] and while the protocols are not
+identical, it may be reasonable to allow this data to serve as 
+informed priors for the eCPR parameter, which can be updated with the INCEPTION data.
 
-$$\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}$$ where
-$\alpha$ and $\beta$ are the number of successes and failures respectively.
-The difference between the two treatment arms of these previous studies can then serve as the prior information which in the Bayesian paradigm can be updated with the INCEPTION data to form a posterior distribution.
+This prior
+information of the probability of eCPR success in each previous trial, $X_i$, can be summarized as a normal distribution with a mean equal to the proportion of successes, $\hat{p_i}$ with a standard deviation equal to $$\sqrt{\hat{p_i}*(1-\hat{p_i})}$$
+As  baseline success rates for standard CPR varies markedly between the three studies, it was decided to maintain the INCEPTION control baseline with the vague prior for all analyses and to update only the eCPR arm with prior information.
 
@@
 [@mcelreath2020]. We not only calculated the posterior probability of
-any benefit (OR \>1.00), but also of clinically meaningful benefits
+any additional survival with eCPR (OR \>1.00), but also of clinically meaningful benefits
 (defined as OR \>1.10).
 
-All analyses were executed within the integrated development environment
-of RStudio and the statistical code can be found on Github
+ITT assesses subjects based on the group they were initially (and randomly) allocated
+to, regardless of whether or not they dropped out, were fully adhered to
+the treatment or switched to an alternative treatment. ITT analyses can
+therefore be seen as a conservative estimate which mirrors clinical
+effectiveness. In contrast, a per protocol (PP) analysis involves a
+comparison of treatment groups in a trial that includes only those
+patients who completed the treatment they were originally allocated to.
+Similarly an "as treated" analysis considers only which treatments
+subjects received, regardless of their randomization status and protocol
+adherence. While both PP and as-treated analyses alone may lead to bias,
+in conjunction with an ITT analysis they may provide additional insights
+into efficacy and have also been examined from a Bayesian perspective.
+
+All analyses were executed using `R`[@R]  within the integrated development environment
+of RStudio. Bayesian analyses were performed using the front end `rstanarm` package[@rstanarm] to the `Stan` programming language[stan]. The statistical code can be found on Github
 (https://github.com/brophyj/eCPR).
@@
 # Results
-
-The ITT data from the three pertinent trials is shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using a vague prior, produces an odds ratio 1.32 (95% Credible Interval (CrI) 0.54 - 3.22).
-The closeness of this result to the original analysis confirms the lack if impact of the vague prior and shows this Bayesian analysis is dominated by the observed data.
-
-\newpage
-
-```{r echo=FALSE,  message=FALSE, warning=FALSE, include=FALSE, results='hide'}
-knitr::opts_chunk$set(echo = FALSE)
-options(scipen = 1, digits = 3)
-pander::panderOptions('keep.trailing.zeros', TRUE)
-
-library(here)
-library(tidyverse)
-library(broom)
-library(cobalt)
-library(table1)
-library(knitr)
-library(patchwork)
-library(rstanarm)
-library(tidybayes)
-
-```
-
-# Tables
-
-```{r}
+```{r cache=TRUE}
 df <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
@@
 df1 <- df[, c(1:4)]  %>% tidyr::pivot_wider(names_from = Tx, values_from = c(fail,success))
-
-knitr::kable(df1, caption = "Table 1 Extracted ITT trial data", col.names=c("Trial", "Fail CPR", "Fail eCPR", "Success CPR", "Success eCPR"))
-
 #frequentist INCEPTION  to reproduce NEJM result
@@
 # [1] 1.300 0.531 3.183
-```
 
-# Bayesian analysis
-
-```{r cache=TRUE}
 # - INCEPTION alone default vague priors
 set.seed(1234)
-fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0)
+fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
+fit1_sum <- summary(fit1, digits=3)
+conf_95 <- c(exp(fit1_sum[2]),exp(fit1_sum[2]+c(-1,1)*1.96*fit1_sum[10]))
+tt <- fit1 %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
 # summary(fit1, digits=3)
@@
 # help('prior_summary.stanreg') 
-
-# informative priors using 2 previous studies
-# one method
-t1 <- rbeta(100000,25,124)
-t2 <- rbeta(100000,45,95)
-t3 <- t2 - t1
-# t3 can be approximated by a normal distribution with the following parameters
-# mean(t3)
-# [1] 0.154
-# sd(t3)
-# [1] 0.0498
-
-# second method
-library(LearnBayes)
-quantile1=list(p=.025, x=quantile(t3,0.025))     # 2.5% quantile should be 0.0556 
-quantile2=list(p=.975, x=quantile(t3,0.975))      # 97.5% quantile should be 0..251
-par_beta <- beta.select(quantile1, quantile2)
-m_beta <-  par_beta[1]/(par_beta[1]+par_beta[2])
-v_beta <- (par_beta[1] * par_beta[2])/((par_beta[1]+par_beta[2])^2 * (par_beta[1] +par_beta[2] + 1))
-sd_beta <- sqrt(v_beta)
-# Beta (6.35, 39.54) gives mean = 0.138 sd = 0.0504 close to estimates using method 1
-
-# stan_glm doesn't seem able to take a Beta prior distribution so will use normal approximation
-
-fit2 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0,
-                 prior = normal(mean(t3), sd(t3)))
-summary(fit2, digits=4)
+fig_prob_1.1 <- sum(tt$TxeCPR>0.0953)/nrow(tt) # OR >1.1 (log (1.1) = 0.0953)
+# [1] 0.657
 
@@
 
-**Have obviously not done this correctly as the prior should have shifted the default model parameter for eCPR to a larger number with a small sd**
+ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6751;@RN6759] that also randomized CPR to eCPR are shown in Table 1.
+Performing a Bayesian analysis on the INCEPTION trial, using the default vague prior, produces an odds ratio (OR) 1.32 (95% Credible Interval (CrI) 0.54 - 3.22).
+The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms no impact of the vague prior and shows this Bayesian analysis is completely dominated by the observed INCEPTION data.
+
+One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved survival with eCPR.
+The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,2)`%.
+The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1,2)`%.
+
+```{r priors, cache=TRUE}
+## running using prior for eCPR from different studies but leaving CPR with vague prior and INCEPTION data since baseline risks with CPR seem 
+## to vary widely btw the different studies
+
+eCRP_bayes_no_intercept_prior <- function(data, m_eCPR, sd_eCPR ) {
+  set.seed(1234)
+  fit <- stan_glm(prop_success ~ Tx, data = data, family = binomial(link="logit"), weights = total, refresh=0,
+                  prior = normal(m_eCPR, sd_eCPR))
+  fit_sum <- summary(fit, digits=3)
+  conf_95 <- c(exp(fit_sum[2]),exp(fit_sum[2]+c(-1,1)*1.96*fit_sum[10]))
+  tt <- fit %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+  fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
+  fit_draws <- fit %>% tidybayes::tidy_draws() 
+  my_list <- list(fig_prob, conf_95, fit_draws)
+  return(my_list)
+}
+
+# combined_prior 
+# informative priors using 2 previous studies combined
+comb_m_CPR <- sum(df1[c(2,3),4]) / (sum(df1[c(2,3),4]) + sum(df1[c(2,3),2])) # proportion previous results for CPR
+comb_sd_CPR <- sqrt(comb_m_CPR*(1-comb_m_CPR))                               
+comb_m_eCPR <- sum(df1[c(2,3),5]) / (sum(df1[c(2,3),5]) + sum(df1[c(2,3),3])) # proportion previous results for eCPR
+comb_sd_eCPR <- sqrt(comb_m_eCPR*(1-comb_m_eCPR)) #sd
+
+# enthusiastic prior with only ARREST study
+enth_m_CPR <- sum(df1[c(2),4]) + 1 / (sum(df1[c(2),4]) + sum(df1[c(2),2])) # proportion previous results for CPR, + 0.5 to avoid 0 cell
+enth_sd_CPR <- sqrt(enth_m_CPR*(1-enth_m_CPR))
+enth_m_eCPR <- sum(df1[c(2),5]) / (sum(df1[c(2),5]) + sum(df1[c(2),3])) # proportion previous results for eCPR
+enth_sd_eCPR <- sqrt(enth_m_eCPR*(1-enth_m_eCPR)) # sd
+
+# skeptical prior with only PRAGUE study
+skep_m_CPR <- sum(df1[c(3),4])  / (sum(df1[c(3),4]) + sum(df1[c(3),2])) # proportion previous results for CPR
+skep_sd_CPR <- sqrt(skep_m_CPR*(1-skep_m_CPR))
+skep_m_eCPR <- sum(df1[c(3),5]) / (sum(df1[c(3),5]) + sum(df1[c(3),3])) # proportion previous results for eCPR
+skep_sd_eCPR <- sqrt(skep_m_eCPR*(1-skep_m_eCPR)) #sd
+
+comb_prior1 <-eCRP_bayes_no_intercept_prior(df[c(1:2),], comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
+enth_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
+skep_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior
+
+# >OR 1.1
+comb_prob_1.1 <- sum(comb_prior1[[3]]$TxeCPR > 0.0953) / nrow(comb_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.732
+enth_prob_1.1 <- sum(enth_prior1[[3]]$TxeCPR > 0.0953) / nrow(enth_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.77
+skep_prob_1.1 <- sum(skep_prior1[[3]]$TxeCPR > 0.0953) / nrow(skep_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.71
+
+```
+
+Another advantage of a Bayesian approach is the possibility of including previous evidence as informed priors.
+We considered three possible scenarios i) using all the available prior RCT data[@RN6751;@RN6759] to create a combined prior which was expressed as a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`)
+ii) using only the ARREST data, labelled an enthusiastic prior, as this trial was stopped prematurely for efficacy, expressed as a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`)    
+iii) using only the PRAGUE data, labelled a skeptical prior, as this trial was stopped prematurely for futility, expressed as a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`)
+
+These different prior probabilities can be updated using the INCEPTION to create posterior distributions that are summarized in Table 2.
+Basically, it is observed that in all cases the point estimate for the enhanced eCPR survival remains fairly constant but the associated uncertainty is reduced with the additional data as reflected by the narrower 95% CrI.
+The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.
+
+\newpage
+
+# Tables
+
+```{r}
+# knitr::kable(df1, caption = "Table 1 Extracted ITT trial data", col.names=c("Trial", "Fail CPR", "Fail eCPR", "Success CPR", "Success eCPR"))
+
+gt(df1) %>%
+  tab_header(
+    title = "Table 1 Extracted ITT trial data"
+  ) %>%
+  cols_label(
+    Trial = html("Trial"),
+    fail_CPR = html("Fail CPR (n)"),
+    fail_eCPR = html("Fail eCPR (n)"),
+    success_CPR = html("Success CPR (n)"),
+    success_eCPR = html("Success eCPR (n)"),
+  ) %>%
+  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
+  cols_align(
+  align = c("center"),
+  columns = c(fail_CPR, fail_eCPR, success_CPR, success_eCPR )
+  ) %>%
+  tab_source_note(
+    source_note = "eCPR = extracorporeal cardiopulmonary resuscitation"
+  )
+
+table_tbl_2 <- 
+  tibble(
+    prior = c("Vague", "Combined", "Enthusiastic", "Skeptical"),
+    pt_est = c(conf_95[1], comb_prior1[[2]][1], enth_prior1[[2]][1], skep_prior1[[2]][1]),
+    llimit = c(conf_95[2], comb_prior1[[2]][2], enth_prior1[[2]][2], skep_prior1[[2]][2]),
+    ulimit = c(conf_95[3], comb_prior1[[2]][3], enth_prior1[[2]][3], skep_prior1[[2]][3]),
+    p.gt.1 = c(fig_prob, comb_prior1[1], enth_prior1[1], skep_prior1[1])
+  )
+
+gt(table_tbl_2) %>%
+  tab_header(
+    title = "Table 2 Probabilities of improved survival with eCPR with various priors"
+  ) %>%
+  tab_source_note(
+    source_note = "Vague: default vague prior"
+  ) %>%
+  tab_source_note(
+    source_note = "Combined: prior eCPR data from ARREST + PRAGUE"
+  ) %>%
+  tab_source_note(
+    source_note = md("Enthusiastic: prior eCPR data from ARREST alone")
+  ) %>%
+  tab_source_note(
+    source_note = md("Skeptical: prior eCPR data from PRAGUE alone")
+  ) %>%
+  tab_spanner(
+    label = "Priors",
+    columns = c(prior)
+   ) %>%
+  tab_spanner(
+    label = "OR",
+    columns = c(pt_est)
+  ) %>%
+  tab_spanner(
+    label = "Probability",
+    columns = c(p.gt.1)
+   ) %>%
+  tab_spanner(
+    label = "95% CrI",
+    columns = c(llimit, ulimit)
+  ) %>%
+  cols_label(
+    prior = html(""),
+    pt_est = html("point estimate"),
+    llimit = html("lower limit"),
+    ulimit = html("upper limit"),
+    p.gt.1 = html("p(OR) >1 ")
+  ) %>%
+  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
+  cols_align(
+  align = c("center"),
+  columns = c(pt_est, llimit, ulimit)
+)
+
+```
+
+\newpage
 
@@
 
-```{r}
-
-# tt <- fit1 %>%
-#  spread_draws(`(Intercept)`, `TxeCPR`)
-# sum(tt$TxeCPR>0)/nrow(tt)
+```{r fig1, warning=FALSE, message=FALSE, fig.width=8}
 
@@
   scale_y_continuous(name="Density") +
-  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior") +
-  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 71%") +
+  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior", caption = "AUC to the right of vertical line at HR =1.1 = 65.7%") +
+  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .90)) +
   theme_classic() +
@@
 
-# Examining prior information
-x <- seq(0, .5, by = 0.001)
+```
 
-#THEN EACH OF OUR BETAS DISTR
+\newpage
 
-beta1 <- dbeta(x, shape1 = 25, shape2 = 124) # probability of success with CPR
-beta2 <- dbeta(x, shape1 = 45, shape2 = 95)  # probability of success with eCPR
-beta3 <- dbeta(x, shape1 = 6.36, shape2 = 39.65)
+```{r fig2, warning=FALSE, message=FALSE, fig.height=15, fig.width=8}
+g1 <- comb_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
+  ggplot(aes(x = exp(TxeCPR),
+             fill = after_stat(x > 1),
+             slab_alpha = after_stat((x)))
+  ) +
+  ggdist::stat_halfeye() +
+  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_y_continuous(name="Density") +
+  labs(title = "Figure 2a INCEPTION ITT analysis with combined prior*", subtitle = "*data from ARREST and PRAGUE",
+       caption = "AUC to the right of vertical line at HR =1.1 = 73.2%") +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 81.2% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .86)) +
+  theme_classic() +
+  theme(legend.position="none")
 
-mixture <- ggplot()+
-  geom_line(aes(x, beta1),  color = "red",
-            linewidth = 1.2) +
-  geom_line(aes(x, beta2), color = "green",
-            linewidth = 1.2) +
-  geom_line(aes(x, beta3), color = "black",
-            linewidth = 1.2) +
-  labs(title = "Figure 2 Beta Distribution of prior studies",
-       x = "Sequence",
-       y = "Beta",
-       caption = "Green = p(+outcome | eCPR) \nRed = p(+outcome | CPR) \nBlack= difference of the 2 probabilities")
+g2 <- enth_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
+  ggplot(aes(x = exp(TxeCPR),
+             fill = after_stat(x > 1),
+             slab_alpha = after_stat((x)))
+  ) +
+  ggdist::stat_halfeye() +
+  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_y_continuous(name="Density") +
+  labs(title = "Figure 2b INCEPTION ITT analysis with enthusiastic prior*", subtitle = "*data from ARREST alone",
+       caption = "AUC to the right of vertical line at HR =1.1 = 77.0%") +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 84.9% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .77)) +
+  theme_classic() +
+  theme(legend.position="none")
 
-mixture
+g3 <- skep_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
+  ggplot(aes(x = exp(TxeCPR),
+             fill = after_stat(x > 1),
+             slab_alpha = after_stat((x)))
+  ) +
+  ggdist::stat_halfeye() +
+  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_y_continuous(name="Density") +
+  labs(title = "Figure 2c INCEPTION ITT analysis with skeptical prior*", subtitle = "*data from PRAGUE alone",
+       caption = "AUC to the right of vertical line at HR =1.1 = 71.0%") +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 80.4% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .88)) +
+  theme_classic() +
+  theme(legend.position="none")
+
+g1 / g2 / g3 +
+  plot_annotation(title = 'Figure 2. Probability density plots with informative priors')
 

### 2023-02-08 manuscript/manuscript.tex
updated Tables and Graphs

--- 2023-02-07 manuscript/manuscript.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 extracorporeal CPR (the addition of extracorporeal membrane oxygenation
-to standard advanced cardiac life support) can improve survival and
-diminish anoxic brain injury is a current topic of research. The largest
-randomized clinical trial (RCT) examining this question recently
+to standard advanced cardiac life support (eCPR)) can improve survival
+and diminish anoxic brain injury is a current topic of research. The
+largest randomized clinical trial (RCT) examining this question recently
 published their results\citep{CPR2023a}. For the primary outcome, 30 day
@@
 randomization as well as assuring the validity of the statistical
-analyses. ITT assesses subjects based on the group they were initially
-(and randomly) allocated to, regardless of whether or not they dropped
-out, were fully adhered to the treatment or switched to an alternative
-treatment. ITT analyses can therefore be seen as a conservative estimate
-which mirrors clinical effectiveness. In contrast, a per protocol (PP)
-analysis involves a comparison of treatment groups in a trial that
-includes only those patients who completed the treatment they were
-originally allocated to. Similarly an ``as treated'' analysis considers
-only which treatments subjects received, regardless of their
-randomization status and protocol adherence. While both PP and
-as-treated analyses alone may lead to bias, in conjunction with an ITT
-analysis they may provide additional insights into efficacy and have
-also been examined from a Bayesian perspective.
+analyses.
 
@@
 analyses to learn sequentially. There were two previous RCTs examining
-extracorporeal CPR{[}\citet{RN6759}{]}\citep{RN6751} and while the
-protocols are not identical, it may be reasonable to allow this data can
-serve as an informed prior which can be updated with the INCEPTION data.
-This prior information of the probability of success in each arm,
-\(X_i\), can be summarized by a Beta distribution as follow
-\[X_i\sim\mbox{Beta}(\alpha_0,\beta_0)\] where in the context of
-Bernoulli trials, \(\alpha\) can be interpreted as 1 + number of
-successes and \(\beta\) can be interpreted as 1 + number of failures.
-The mean and variance of this distribution are
-\[\mu=\frac{\alpha}{\alpha+\beta}\]
+extracorporeal CPR\citep{RN6759, RN6751} and while the protocols are not
+identical, it may be reasonable to allow this data to serve as informed
+priors for the eCPR parameter, which can be updated with the INCEPTION
+data.
 
-\[\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}\] where
-\(\alpha\) and \(\beta\) are the number of successes and failures
-respectively. The difference between the two treatment arms of these
-previous studies can then serve as the prior information which in the
-Bayesian paradigm can be updated with the INCEPTION data to form a
-posterior distribution.
+This prior information of the probability of eCPR success in each
+previous trial, \(X_i\), can be summarized as a normal distribution with
+a mean equal to the proportion of successes, \(\hat{p_i}\) with a
+standard deviation equal to \[\sqrt{\hat{p_i}*(1-\hat{p_i})}\] As
+baseline success rates for standard CPR varies markedly between the
+three studies, it was decided to maintain the INCEPTION control baseline
+with the vague prior for all analyses and to update only the eCPR arm
+with prior information.
 
@@
 \citep{mcelreath2020}. We not only calculated the posterior probability
-of any benefit (OR \textgreater1.00), but also of clinically meaningful
-benefits (defined as OR \textgreater1.10).
+of any additional survival with eCPR (OR \textgreater1.00), but also of
+clinically meaningful benefits (defined as OR \textgreater1.10).
 
-All analyses were executed within the integrated development environment
-of RStudio and the statistical code can be found on Github
-(https://github.com/brophyj/eCPR).
+ITT assesses subjects based on the group they were initially (and
+randomly) allocated to, regardless of whether or not they dropped out,
+were fully adhered to the treatment or switched to an alternative
+treatment. ITT analyses can therefore be seen as a conservative estimate
+which mirrors clinical effectiveness. In contrast, a per protocol (PP)
+analysis involves a comparison of treatment groups in a trial that
+includes only those patients who completed the treatment they were
+originally allocated to. Similarly an ``as treated'' analysis considers
+only which treatments subjects received, regardless of their
+randomization status and protocol adherence. While both PP and
+as-treated analyses alone may lead to bias, in conjunction with an ITT
+analysis they may provide additional insights into efficacy and have
+also been examined from a Bayesian perspective.
+
+All analyses were executed using \texttt{R}\citep{R} within the
+integrated development environment of RStudio. Bayesian analyses were
+performed using the front end \texttt{rstanarm} package\citep{rstanarm}
+to the \texttt{Stan} programming language{[}stan{]}. The statistical
+code can be found on Github (https://github.com/brophyj/eCPR).
 
@@
 
-The ITT data from the three pertinent trials is shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using a vague
-prior, produces an odds ratio 1.32 (95\% Credible Interval (CrI) 0.54 -
-3.22). The closeness of this result to the original analysis confirms
-the lack if impact of the vague prior and shows this Bayesian analysis
-is dominated by the observed data.
+ITT data from the INCEPTION trial\citep{CPR2023a} and two other
+pertinent trials\citep{RN6751, RN6759} that also randomized CPR to eCPR
+are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
+trial, using the default vague prior, produces an odds ratio (OR) 1.32
+(95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
+to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms no impact
+of the vague prior and shows this Bayesian analysis is completely
+dominated by the observed INCEPTION data.
+
+One of the advantages of a Bayesian approach is the ability to make
+direct probability statements about the estimand of improved survival
+with eCPR. The eCPR probability density function for improved survival
+from INCEPTION data with the default vague prior is displayed in Figure
+1 and reveals that the probability of enhanced survival with eCPR is
+72.72\%. The probability that the improved survival exceeds a 10\%
+improvement is 0.66\%.
+
+Another advantage of a Bayesian approach is the possibility of including
+previous evidence as informed priors. We considered three possible
+scenarios i) using all the available prior RCT
+data\citep{RN6751, RN6759} to create a combined prior which was
+expressed as a N(0.32, 0.47)\\
+ii) using only the ARREST data, labelled an enthusiastic prior, as this
+trial was stopped prematurely for efficacy, expressed as a N(0.43,
+0.49)\\
+iii) using only the PRAGUE data, labelled a skeptical prior, as this
+trial was stopped prematurely for futility, expressed as a N(0.31, 0.46)
+
+These different prior probabilities can be updated using the INCEPTION
+to create posterior distributions that are summarized in Table 2.
+Basically, it is observed that in all cases the point estimate for the
+enhanced eCPR survival remains fairly constant but the associated
+uncertainty is reduced with the additional data as reflected by the
+narrower 95\% CrI. The posterior probability of improved eCPR survival
+goes from 80\% with a skeptical prior to 85\% with the enthusiastic
+prior.
 
@@
 
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{Table 1 Extracted ITT trial data}\tabularnewline
-\toprule()
-Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
-\midrule()
-\endfirsthead
-\toprule()
-Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
-\midrule()
-\endhead
+\setlength{\LTpost}{0mm}
+\begin{longtable}{lcccc}
+\caption*{
+{\large Table 1 Extracted ITT trial data}
+} \\ 
+\toprule
+Trial & Fail CPR (n) & Fail eCPR (n) & Success CPR (n) & Success eCPR (n) \\ 
+\midrule
 INCEPTION & 52 & 56 & 10 & 14 \\ 
@@
 PRAGUE & 108 & 86 & 24 & 38 \\ 
-\bottomrule()
-\end{longtable}
-
-\hypertarget{bayesian-analysis}{%
-\section{Bayesian analysis}\label{bayesian-analysis}}
-
-\begin{longtable}{lrrrc}
-\toprule
-prior & pt\_est <- ... & llimit <- ... & ulimit <- ... & p.gt.1 <- ... \\ 
-\midrule
-vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
-combined & 1.27 & 0.696 & 2.31 & 0.782 \\ 
-enthusiastic & 1.27 & 0.708 & 2.27 & 0.79 \\ 
-skeptical & 1.25 & 0.695 & 2.26 & 0.777 \\ 
 \bottomrule
 \end{longtable}
+\begin{minipage}{\linewidth}
+eCPR = extracorporeal cardiopulmonary resuscitation\\
+\end{minipage}
 
-\hypertarget{summary-of-what-i-dont-understand}{%
-\section{\texorpdfstring{\textbf{Summary of what I don't
-understand}}{Summary of what I don't understand}}\label{summary-of-what-i-dont-understand}}
+\setlength{\LTpost}{0mm}
+\begin{longtable}{lcccc}
+\caption*{
+{\large Table 2 Probabilities of improved survival with eCPR with various priors}
+} \\ 
+\toprule
+Priors & OR & \multicolumn{2}{c}{95\% CrI} & Probability \\ 
+\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-5}
+ & point estimate & lower limit & upper limit & p(OR) >1  \\ 
+\midrule
+Vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
+Combined & 1.35 & 0.705 & 2.58 & 0.817 \\ 
+Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 \\ 
+Skeptical & 1.32 & 0.700 & 2.50 & 0.804 \\ 
+\bottomrule
+\end{longtable}
+\begin{minipage}{\linewidth}
+Vague: default vague prior\\
+Combined: prior eCPR data from ARREST + PRAGUE\\
+Enthusiastic: prior eCPR data from ARREST alone\\
+Skeptical: prior eCPR data from PRAGUE alone\\
+\end{minipage}
 
-This makes only partial sense. 95\% CrI are narrower with informative
-priors. OK, but one would think the enthusiastic prior from a trial
-ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes
-for eCPR vs CPR would shift the HR of 1.32 to the right but instead the
-HR falls to have a 1.27. The skeptical prior from PRAGUE which was
-stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as
-expected
-
-So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)\\
-skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)
-
-Tx for enthusiastic prior \textgreater{} Tx skeptical prior (eCPR) but
-this seems to be offset by the reverse occurring for the intercept
-(CPR).\\
-The problem seems to be in how comparable are these previous RCTs as the
-baseline success for the standard TX (CRP) is radically different 0/14
-vs 24/132.
-
-When I use a prior that uses the enthusiastic intercept prior and the
-skeptical Tx prior, I get a bigger downward shift with HR 1.220 95\%
-CrI(0.678 2.194) and P\textgreater1 74.5\%
-
-Can you see any other explanation?
-
-Do you have any other insights?
+\newpage
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/fig1-1.pdf}
+
+\newpage
+
+\includegraphics{manuscript_files/figure-pdf/fig2-1.pdf}
 

### 2023-02-10 manuscript/manuscript.qmd
ROPE

@@
 # help('prior_summary.stanreg') 
-fig_prob_1.1 <- sum(tt$TxeCPR>0.0953)/nrow(tt) # OR >1.1 (log (1.1) = 0.0953)
+# P > 1.1
+fig_prob_1.1 <- sum(tt$TxeCPR > 0.0953)/nrow(tt) # OR >1.1 (log (1.1) = 0.0953)
 # [1] 0.657
 
+# P > 1/1.1
+fig_prob_1.1.1 <- sum(tt$TxeCPR > -0.0954)/nrow(tt) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.791
+
+# difference equivalence
+
+vague_eq <- fig_prob_1.1.1 - fig_prob_1.1
 ```
@@
 
+# > OR 1/ 1.1 (0.909)
+comb_prob_1.1.1 <- sum(comb_prior1[[3]]$TxeCPR > -0.0954) / nrow(comb_prior1[[3]]) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.884
+enth_prob_1.1.1 <- sum(enth_prior1[[3]]$TxeCPR > -0.0954) / nrow(enth_prior1[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.908
+skep_prob_1.1.1 <- sum(skep_prior1[[3]]$TxeCPR > -0.0954) / nrow(skep_prior1[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.878
+
+# equivalence +/- 10%
+comb_eq <- comb_prob_1.1.1 - comb_prob_1.1 
+enth_eq <- enth_prob_1.1.1 - enth_prob_1.1
+skep_eq <- skep_prob_1.1.1 - skep_prob_1.1
 ```
@@
 These different prior probabilities can be updated using the INCEPTION to create posterior distributions that are summarized in Table 2.
-Basically, it is observed that in all cases the point estimate for the enhanced eCPR survival remains fairly constant but the associated uncertainty is reduced with the additional data as reflected by the narrower 95% CrI.
+Basically, in all cases the point estimate for the enhanced eCPR survival remains fairly constant but the associated uncertainty is reduced with the additional data as reflected by the narrower 95% CrI.
 The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.
+The graphical presentation of these results is shown in Figure 2.
+
+Although INCEPTION did not report a per-protocol analysis, just an analysis may be revealing concerning treatment efficacy, at the risk of an increased risk of bias by not respecting the ITT.
+From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the CPR group compared to 5 survivors from 46 patients receiving eCPR.
+With a vague prior, the probability of increased survival with eCPR is decreased compared to CPR but with wide CrI (OR 0.45, 95% CrI 1.5 - 1.35), limiting any definitive conclusions.
+The probabilities of increased eCPR survival incorporating the previous identified priors is shown in Table 3.
+
+```{r PP, cache=TRUE}
+# per protocol - only data for INCEPTION, other data just placeholder if PP data becomes available
+df_pp <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
+             fail = c(48, 41, 15, 8, 108, 86),
+             success = c(13,5, 0, 6,24,38 )) %>% 
+  mutate(total = fail + success,
+         prop_success = success / total)
+
+# df_pp[c(1:2),] data set used for PP analyses
+
+set.seed(1234)
+fit1_pp <- stan_glm(prop_success ~ Tx, data = df_pp[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
+fit1_pp_sum <- summary(fit1_pp, digits=3)
+conf_95_pp <- c(exp(fit1_pp_sum[2]),exp(fit1_pp_sum[2]+c(-1,1)*1.96*fit1_pp_sum[10]))
+tt_pp <- fit1_pp %>%
+    spread_draws(`(Intercept)`, `TxeCPR`)
+fig_prob_pp <- sum(tt_pp$TxeCPR>0)/nrow(tt_pp)
+# summary(fit1, digits=3)
+# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
+# get_variables(fit1)
+# prior_summary(fit1)
+# help('prior_summary.stanreg') 
+fig_prob_1.1_pp <- sum(tt_pp$TxeCPR>0.0953)/nrow(tt_pp) #
+fig_prob_1.1.1_pp <- sum(tt_pp$TxeCPR>-0.0954)/nrow(tt_pp)
+
+comb_prior1_pp <-eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
+enth_prior1_pp <- eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
+skep_prior1_pp <- eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior
+
+# >OR 1.1
+comb_prob_pp_1.1 <- sum(comb_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(comb_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.732
+enth_prob_pp_1.1 <- sum(enth_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(enth_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.77
+skep_prob_pp_1.1 <- sum(skep_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(skep_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.71
+
+# > OR 1/ 1.1 (0.909)
+comb_prob_pp_1.1.1 <- sum(comb_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(comb_prior1_pp[[3]]) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.884
+enth_prob_pp_1.1.1 <- sum(enth_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(enth_prior1_pp[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.908
+skep_prob_pp_1.1.1 <- sum(skep_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(skep_prior1_pp[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.878
+
+# equivalence +/- 10%
+vague_eq_pp <- fig_prob_1.1.1_pp - fig_prob_1.1_pp
+comb_eq_pp <- comb_prob_pp_1.1.1 - comb_prob_pp_1.1
+enth_eq_pp <- enth_prob_pp_1.1.1 - enth_prob_pp_1.1
+skep_eq_pp <- skep_prob_pp_1.1.1 - skep_prob_pp_1.1
+
+```
 
@@
 
+## Table 2
+
 table_tbl_2 <- 
@@
     ulimit = c(conf_95[3], comb_prior1[[2]][3], enth_prior1[[2]][3], skep_prior1[[2]][3]),
-    p.gt.1 = c(fig_prob, comb_prior1[1], enth_prior1[1], skep_prior1[1])
+    p.gt.1 = c(fig_prob, comb_prior1[1], enth_prior1[1], skep_prior1[1]),
+    p.eq   = c(vague_eq, comb_eq, enth_eq, skep_eq)
   )
@@
   tab_header(
-    title = "Table 2 Probabilities of improved survival with eCPR with various priors"
+    title = "Table 2 eCPR odds ratios, 95% credible intervals and probabilities with various priors"
   ) %>%
@@
   ) %>%
+  tab_source_note(
+    source_note = md("ROPE: range of practical equivalence = + / - 10% OR (odds ratio)")
+  ) %>%
   tab_spanner(
@@
   tab_spanner(
-    label = "Probability",
-    columns = c(p.gt.1)
+    label = "Probabilities",
+    columns = c(p.gt.1, p.eq)
    ) %>%
@@
     ulimit = html("upper limit"),
-    p.gt.1 = html("p(OR) >1 ")
+    p.gt.1 = html("p(OR) >1 "),
+    p.eq   = html(" p(ROPE)")
+  ) %>%
+  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
+  cols_align(
+  align = c("center"),
+  columns = c(pt_est, llimit, ulimit)
+)
+
+## Table 3
+
+table_tbl_3 <- 
+  tibble(
+    prior = c("Vague", "Combined", "Enthusiastic", "Skeptical"),
+    pt_est = c(conf_95_pp[1], comb_prior1_pp[[2]][1], enth_prior1_pp[[2]][1], skep_prior1_pp[[2]][1]),
+    llimit = c(conf_95_pp[2], comb_prior1_pp[[2]][2], enth_prior1_pp[[2]][2], skep_prior1_pp[[2]][2]),
+    ulimit = c(conf_95_pp[3], comb_prior1_pp[[2]][3], enth_prior1_pp[[2]][3], skep_prior1_pp[[2]][3]),
+    p.gt.1 = c(fig_prob_pp, comb_prior1_pp[1], enth_prior1_pp[1], skep_prior1_pp[1]),
+    p.eq   = c(vague_eq_pp, comb_eq_pp, enth_eq_pp, skep_eq_pp)
+  )
+
+gt(table_tbl_3) %>%
+  tab_header(
+    title = "Table 3 eCPR (per protocol) odds ratios, 95% credible intervals and probabilities with various priors"
+  ) %>%
+  tab_source_note(
+    source_note = "Vague: default vague prior"
+  ) %>%
+  tab_source_note(
+    source_note = "Combined: prior eCPR data from ARREST + PRAGUE"
+  ) %>%
+  tab_source_note(
+    source_note = md("Enthusiastic: prior eCPR data from ARREST alone")
+  ) %>%
+  tab_source_note(
+    source_note = md("Skeptical: prior eCPR data from PRAGUE alone")
+  ) %>%
+  tab_source_note(
+    source_note = md("ROPE: range of practical equivalence = + / - 10% OR (odds ratio)")
+  ) %>%
+  tab_spanner(
+    label = "Priors",
+    columns = c(prior)
+   ) %>%
+  tab_spanner(
+    label = "OR",
+    columns = c(pt_est)
+  ) %>%
+  tab_spanner(
+    label = "Probabilities",
+    columns = c(p.gt.1, p.eq)
+   ) %>%
+  tab_spanner(
+    label = "95% CrI",
+    columns = c(llimit, ulimit)
+  ) %>%
+  cols_label(
+    prior = html(""),
+    pt_est = html("point estimate"),
+    llimit = html("lower limit"),
+    ulimit = html("upper limit"),
+    p.gt.1 = html("p(OR) >1 "),
+    p.eq   = html(" p(ROPE)")
   ) %>%
@@
   ggdist::stat_halfeye() +
-  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
   scale_y_continuous(name="Density") +
   labs(title = "Figure 1 INCEPTION ITT analysis with vague prior", caption = "AUC to the right of vertical line at HR =1.1 = 65.7%") +
-  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
   geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .90)) +
+  geom_segment(aes(x = 1 / 1.1, y = 0, xend = 1 / 1.1, yend = .86)) +
+  annotate(geom="text", x=2.0,y=.45, label = "Probability (range) of practical equivalence (+/- 10% OR) \nAUC between 2 black vertical lines =13.4% ", hjust = 0) +
   theme_classic() +
@@
   ggdist::stat_halfeye() +
-  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
+  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
   scale_y_continuous(name="Density") +

### 2023-02-10 manuscript/manuscript.tex
ROPE

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 to create posterior distributions that are summarized in Table 2.
-Basically, it is observed that in all cases the point estimate for the
-enhanced eCPR survival remains fairly constant but the associated
-uncertainty is reduced with the additional data as reflected by the
-narrower 95\% CrI. The posterior probability of improved eCPR survival
-goes from 80\% with a skeptical prior to 85\% with the enthusiastic
-prior.
+Basically, in all cases the point estimate for the enhanced eCPR
+survival remains fairly constant but the associated uncertainty is
+reduced with the additional data as reflected by the narrower 95\% CrI.
+The posterior probability of improved eCPR survival goes from 80\% with
+a skeptical prior to 85\% with the enthusiastic prior. The graphical
+presentation of these results is shown in Figure 2.
+
+Although INCEPTION did not report a per-protocol analysis, just an
+analysis may be revealing concerning treatment efficacy, at the risk of
+an increased risk of bias by not respecting the ITT. From INCEPTION
+Figure S4\citep{CPR2023a}, it appears that the per protocol data for
+mortality is 13 survivors from 61 patients in the CPR group compared to
+5 survivors from 46 patients receiving eCPR. With a vague prior, the
+probability of increased survival with eCPR is decreased compared to CPR
+but with wide CrI (OR 0.45, 95\% CrI 1.5 - 1.35), limiting any
+definitive conclusions. The probabilities of increased eCPR survival
+incorporating the previous identified priors is shown in Table 3.
 
@@
 \setlength{\LTpost}{0mm}
-\begin{longtable}{lcccc}
+\begin{longtable}{lccccr}
 \caption*{
-{\large Table 2 Probabilities of improved survival with eCPR with various priors}
+{\large Table 2 eCPR odds ratios, 95\% credible intervals and probabilities with various priors}
 } \\ 
 \toprule
-Priors & OR & \multicolumn{2}{c}{95\% CrI} & Probability \\ 
-\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-5}
- & point estimate & lower limit & upper limit & p(OR) >1  \\ 
+Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
+\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
+ & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
 \midrule
-Vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
-Combined & 1.35 & 0.705 & 2.58 & 0.817 \\ 
-Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 \\ 
-Skeptical & 1.32 & 0.700 & 2.50 & 0.804 \\ 
+Vague & 1.32 & 0.543 & 3.21 & 0.727 & 0.134 \\ 
+Combined & 1.35 & 0.705 & 2.58 & 0.817 & 0.153 \\ 
+Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 & 0.138 \\ 
+Skeptical & 1.32 & 0.700 & 2.50 & 0.804 & 0.168 \\ 
 \bottomrule
@@
 Skeptical: prior eCPR data from PRAGUE alone\\
+ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
+\end{minipage}
+
+\setlength{\LTpost}{0mm}
+\begin{longtable}{lccccr}
+\caption*{
+{\large Table 3 eCPR (per protocol) odds ratios, 95\% credible intervals and probabilities with various priors}
+} \\ 
+\toprule
+Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
+\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
+ & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
+\midrule
+Vague & 0.451 & 0.151 & 1.35 & 0.0698 & 0.0473 \\ 
+Combined & 0.859 & 0.430 & 1.71 & 0.33 & 0.2022 \\ 
+Enthusiastic & 0.870 & 0.437 & 1.73 & 0.345 & 0.2013 \\ 
+Skeptical & 0.858 & 0.419 & 1.75 & 0.327 & 0.1970 \\ 
+\bottomrule
+\end{longtable}
+\begin{minipage}{\linewidth}
+Vague: default vague prior\\
+Combined: prior eCPR data from ARREST + PRAGUE\\
+Enthusiastic: prior eCPR data from ARREST alone\\
+Skeptical: prior eCPR data from PRAGUE alone\\
+ROPE: range of practical equivalence = + / - 10\% OR (odds ratio)\\
 \end{minipage}

### 2023-02-10 manuscript/manuscript.qmd
fixed tables 3 decimals, expanded abstract

@@
 ---
-title: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
-subtitle: A Bayesian Perspective
+title: A Bayesian Perspective 
+subtitle: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
 author:
-  - name: James M Brophy
+  - name: James M Brophy MD PhD
     email: james.brophy@mcgill.ca
@@
 abstract: |
-  A recent randomized clinical trial reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR had similar effects on survival with a favorable neurologic outcome.
-Herein, it is examined whether a Bayesian perspective allows any additional insights into the interpretation of this trial.
+  A recent randomized clinical trial, INCEPTION, reported patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome.
+The current study examines if a Bayesian perspective provides  additional insights.
+Depending on the prior selected, the Bayesian approach for the INCEPTION intention-to treat (ITT) analysis shows an equivalence probability between 13.4 - 16.8% (defined as 1 / 1.1 < odds ratio (OR) < 1.1).
+The probability of clinical superiority with eCPR ranges from 65.7 - 77.0 % (defined as OR > 1.1).
+A similar analyses using INCEPTION per protocol data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority of approximately 25%.
+It is concluded that a Bayesian perspective allows considerable additional insights into the trial analysis and interpretation.
+The substantial probabilities of increased survival, but with considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
 keywords: 
@@
 intervals[@RN5420]. Rather it assumes the reader has perhaps heard that
-Bayesian methods mirror our intuitive learning process and is curious
-about its potential application to RCT interpretations.
+Bayesian methods mirror our intuitive learning and diagnostic processes and is curious
+about its potential application to RCT analyses and interpretations.
 
@@
 interval containing 95% of the probability density function
-[@mcelreath2020]. We not only calculated the posterior probability of
-any additional survival with eCPR (OR \>1.00), but also of clinically meaningful benefits
-(defined as OR \>1.10).
+[@mcelreath2020]. Bayesian analyses permit not only calculated the posterior probability of
+any additional survival with eCPR (OR \>1.00), but also of clinically meaningful benefits.
+While there is no universal standard for a clinically meaningful benefit, a survival OR \>1.10 may be an acceptable threshold for many.
+Bayesian analyses also allows calculation of the probability between any two points.
+For example, rather than simply comparing if the survival of one treatment is better, one can calculate a range of practical equivalence (ROPE).
+While different ranges may be proposed as equivalent, +/- 10% seems a reasonable starting point.
 
@@
 therefore be seen as a conservative estimate which mirrors clinical
-effectiveness. In contrast, a per protocol (PP) analysis involves a
-comparison of treatment groups in a trial that includes only those
+effectiveness. In contrast, a per protocol (PP) analysis accounts for adherence by analyzing only those
 patients who completed the treatment they were originally allocated to.
-Similarly an "as treated" analysis considers only which treatments
-subjects received, regardless of their randomization status and protocol
-adherence. While both PP and as-treated analyses alone may lead to bias,
-in conjunction with an ITT analysis they may provide additional insights
-into efficacy and have also been examined from a Bayesian perspective.
+While PP may provide additional insights
+into efficacy, it is subject to bias. Since a PP analysis when performed in conjunction with an ITT analysis may provide additional insights, this has also been subjected to a Bayesian analysis.
 
@@
 ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6751;@RN6759] that also randomized CPR to eCPR are shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using the default vague prior, produces an odds ratio (OR) 1.32 (95% Credible Interval (CrI) 0.54 - 3.22).
-The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms no impact of the vague prior and shows this Bayesian analysis is completely dominated by the observed INCEPTION data.
+Performing a Bayesian analysis on the INCEPTION trial, using the default vague priors, produces an odds ratio (OR) 1.32 (95% Credible Interval (CrI) 0.54 - 3.22).
+The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms minimal impact of the vague prior and shows this Bayesian analysis is completely dominated by the observed INCEPTION data.
 
@@
     p.gt.1 = c(fig_prob, comb_prior1[1], enth_prior1[1], skep_prior1[1]),
+    p.gt.1.1 = c(fig_prob_1.1, as.numeric(comb_prob_1.1), as.numeric(enth_prob_1.1), as.numeric(skep_prob_1.1)),
     p.eq   = c(vague_eq, comb_eq, enth_eq, skep_eq)
@@
     label = "Probabilities",
-    columns = c(p.gt.1, p.eq)
+    columns = c(p.gt.1, p.gt.1.1, p.eq)
    ) %>%
@@
     p.gt.1 = html("p(OR) >1 "),
+    p.gt.1.1 = html("p(OR) >1.1 "),
     p.eq   = html(" p(ROPE)")
@@
   align = c("center"),
-  columns = c(pt_est, llimit, ulimit)
+  columns = c(pt_est, llimit, ulimit, p.gt.1.1, p.eq) 
+  ) %>%
+  fmt_number(
+    columns = c(pt_est, llimit, ulimit, p.gt.1.1),
+    decimals = 3,
+    drop_trailing_zeros = FALSE
 )
@@
     ulimit = c(conf_95_pp[3], comb_prior1_pp[[2]][3], enth_prior1_pp[[2]][3], skep_prior1_pp[[2]][3]),
-    p.gt.1 = c(fig_prob_pp, comb_prior1_pp[1], enth_prior1_pp[1], skep_prior1_pp[1]),
+    p.gt.1 = c(fig_prob_pp, as.numeric(comb_prior1_pp[1]), as.numeric(enth_prior1_pp[1]), as.numeric(skep_prior1_pp[1])),
+    p.gt.1.1 = c(fig_prob_1.1_pp, as.numeric(comb_prob_pp_1.1), as.numeric(enth_prob_pp_1.1), as.numeric(skep_prob_pp_1.1)),
     p.eq   = c(vague_eq_pp, comb_eq_pp, enth_eq_pp, skep_eq_pp)
@@
   align = c("center"),
-  columns = c(pt_est, llimit, ulimit)
+  columns = c(pt_est, llimit, ulimit, p.gt.1, p.gt.1.1, p.eq)
+  ) %>%
+  fmt_number(
+    columns = c(pt_est, llimit, ulimit, p.gt.1, p.gt.1.1, p.eq) ,
+    decimals = 3,
+    drop_trailing_zeros = FALSE
 )
@@
 
-```{r fig1, warning=FALSE, message=FALSE, fig.width=8}
+```{r fig1, warning=FALSE, message=FALSE, fig.height=6}
 
@@
   scale_y_continuous(name="Density") +
-  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior", caption = "AUC to the right of vertical line at HR =1.1 = 65.7%") +
-  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior", caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 65.7%\nthin horizontal line at 0 = 95% CrI with black circle = point estimate\nthick horizontal line = +/- 68% CrI", hjust=0) +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust =0) +
   geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .90)) +
   geom_segment(aes(x = 1 / 1.1, y = 0, xend = 1 / 1.1, yend = .86)) +
-  annotate(geom="text", x=2.0,y=.45, label = "Probability (range) of practical equivalence (+/- 10% OR) \nAUC between 2 black vertical lines =13.4% ", hjust = 0) +
+  annotate(geom="text", x=2.0,y=.35, label = "Probability (range) of practical equivalence (+/- 10% OR) \nAUC between black vertical lines = 13.4% ", hjust = 0) +
   theme_classic() +
@@
 
-```{r fig2, warning=FALSE, message=FALSE, fig.height=15, fig.width=8}
+```{r fig2, warning=FALSE, message=FALSE, fig.height=18, fig.width=8}
 g1 <- comb_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
@@
   labs(title = "Figure 2a INCEPTION ITT analysis with combined prior*", subtitle = "*data from ARREST and PRAGUE",
-       caption = "AUC to the right of vertical line at HR =1.1 = 73.2%") +
+       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 73.2%") +
   annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 81.2% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
@@
   labs(title = "Figure 2b INCEPTION ITT analysis with enthusiastic prior*", subtitle = "*data from ARREST alone",
-       caption = "AUC to the right of vertical line at HR =1.1 = 77.0%") +
+       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 77.0%") +
   annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 84.9% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
@@
   labs(title = "Figure 2c INCEPTION ITT analysis with skeptical prior*", subtitle = "*data from PRAGUE alone",
-       caption = "AUC to the right of vertical line at HR =1.1 = 71.0%") +
+       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 71.0%") +
   annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 80.4% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +

### 2023-02-10 manuscript/manuscript.tex
fixed tables 3 decimals, expanded abstract

@@
 \hypersetup{
-  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
-  pdfauthor={James M Brophy},
+  pdftitle={A Bayesian Perspective},
+  pdfauthor={James M Brophy MD PhD},
   pdfkeywords={extracorporeal CPR, Bayesian statistics},
@@
 \begin{frontmatter}
-\title{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac
-Arrest \\\large{A Bayesian Perspective} }
-\author[1]{James M Brophy%
+\title{A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory
+Out-of-Hospital Cardiac Arrest} }
+\author[1]{James M Brophy MD PhD%
 \corref{cor1}%
@@
 \begin{abstract}
-A recent randomized clinical trial reported in patients with refractory
-out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR
-had similar effects on survival with a favorable neurologic outcome.
-Herein, it is examined whether a Bayesian perspective allows any
-additional insights into the interpretation of this trial.
+A recent randomized clinical trial, INCEPTION, reported patients with
+refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and
+conventional CPR (cCPR) had similar effects on survival with a favorable
+neurologic outcome. The current study examines if a Bayesian perspective
+provides additional insights. Depending on the prior selected, the
+Bayesian approach for the INCEPTION intention-to treat (ITT) analysis
+shows an equivalence probability between 13.4 - 16.8\% (defined as 1 /
+1.1 \textless{} odds ratio (OR) \textless{} 1.1). The probability of
+clinical superiority with eCPR ranges from 65.7 - 77.0 \% (defined as OR
+\textgreater{} 1.1). A similar analyses using INCEPTION per protocol
+data shows an equivalence probability between 4.7 - 20.2\% with reduced
+probabilities of clinical superiority of approximately 25\%. It is
+concluded that a Bayesian perspective allows considerable additional
+insights into the trial analysis and interpretation. The substantial
+probabilities of increased survival, but with considerable residual
+uncertainties, suggests that additional studies are required before
+concluding that eCPR and cCPR have similar survival effects.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 intervals\citep{RN5420}. Rather it assumes the reader has perhaps heard
-that Bayesian methods mirror our intuitive learning process and is
-curious about its potential application to RCT interpretations.
+that Bayesian methods mirror our intuitive learning and diagnostic
+processes and is curious about its potential application to RCT analyses
+and interpretations.
 
@@
 interval containing 95\% of the probability density function
-\citep{mcelreath2020}. We not only calculated the posterior probability
-of any additional survival with eCPR (OR \textgreater1.00), but also of
-clinically meaningful benefits (defined as OR \textgreater1.10).
+\citep{mcelreath2020}. Bayesian analyses permit not only calculated the
+posterior probability of any additional survival with eCPR (OR
+\textgreater1.00), but also of clinically meaningful benefits. While
+there is no universal standard for a clinically meaningful benefit, a
+survival OR \textgreater1.10 may be an acceptable threshold for many.
+Bayesian analyses also allows calculation of the probability between any
+two points. For example, rather than simply comparing if the survival of
+one treatment is better, one can calculate a range of practical
+equivalence (ROPE). While different ranges may be proposed as
+equivalent, +/- 10\% seems a reasonable starting point.
 
@@
 which mirrors clinical effectiveness. In contrast, a per protocol (PP)
-analysis involves a comparison of treatment groups in a trial that
-includes only those patients who completed the treatment they were
-originally allocated to. Similarly an ``as treated'' analysis considers
-only which treatments subjects received, regardless of their
-randomization status and protocol adherence. While both PP and
-as-treated analyses alone may lead to bias, in conjunction with an ITT
-analysis they may provide additional insights into efficacy and have
-also been examined from a Bayesian perspective.
+analysis accounts for adherence by analyzing only those patients who
+completed the treatment they were originally allocated to. While PP may
+provide additional insights into efficacy, it is subject to bias. Since
+a PP analysis when performed in conjunction with an ITT analysis may
+provide additional insights, this has also been subjected to a Bayesian
+analysis.
 
@@
 are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
-trial, using the default vague prior, produces an odds ratio (OR) 1.32
+trial, using the default vague priors, produces an odds ratio (OR) 1.32
 (95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
-to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms no impact
-of the vague prior and shows this Bayesian analysis is completely
+to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms minimal
+impact of the vague prior and shows this Bayesian analysis is completely
 dominated by the observed INCEPTION data.
@@
 \setlength{\LTpost}{0mm}
-\begin{longtable}{lccccr}
+\begin{longtable}{lcccccc}
 \caption*{
@@
 \toprule
-Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
-\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
- & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
+Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{3}{c}{Probabilities} \\ 
+\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-7}
+ & point estimate & lower limit & upper limit & p(OR) >1  & p(OR) >1.1  &  p(ROPE) \\ 
 \midrule
-Vague & 1.32 & 0.543 & 3.21 & 0.727 & 0.134 \\ 
-Combined & 1.35 & 0.705 & 2.58 & 0.817 & 0.153 \\ 
-Enthusiastic & 1.40 & 0.738 & 2.67 & 0.849 & 0.138 \\ 
-Skeptical & 1.32 & 0.700 & 2.50 & 0.804 & 0.168 \\ 
+Vague & $1.321$ & $0.543$ & $3.215$ & 0.727 & $0.657$ & 0.134 \\ 
+Combined & $1.349$ & $0.705$ & $2.580$ & 0.817 & $0.732$ & 0.153 \\ 
+Enthusiastic & $1.403$ & $0.738$ & $2.668$ & 0.849 & $0.770$ & 0.138 \\ 
+Skeptical & $1.322$ & $0.700$ & $2.496$ & 0.804 & $0.710$ & 0.168 \\ 
 \bottomrule
@@
 \toprule
-Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{2}{c}{Probabilities} \\ 
-\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-6}
- & point estimate & lower limit & upper limit & p(OR) >1  &  p(ROPE) \\ 
+Priors & OR & \multicolumn{2}{c}{95\% CrI} & \multicolumn{3}{c}{Probabilities} \\ 
+\cmidrule(lr){1-1} \cmidrule(lr){2-2} \cmidrule(lr){3-4} \cmidrule(lr){5-7}
+ & point estimate & lower limit & upper limit & p(OR) >1  & p(OR) >1.1  &  p(ROPE) \\ 
 \midrule
-Vague & 0.451 & 0.151 & 1.35 & 0.0698 & 0.0473 \\ 
-Combined & 0.859 & 0.430 & 1.71 & 0.33 & 0.2022 \\ 
-Enthusiastic & 0.870 & 0.437 & 1.73 & 0.345 & 0.2013 \\ 
-Skeptical & 0.858 & 0.419 & 1.75 & 0.327 & 0.1970 \\ 
+Vague & $0.451$ & $0.151$ & $1.348$ & $0.070$ & $0.049$ & $0.047$ \\ 
+Combined & $0.859$ & $0.430$ & $1.713$ & $0.330$ & $0.238$ & $0.202$ \\ 
+Enthusiastic & $0.870$ & $0.437$ & $1.734$ & $0.345$ & $0.251$ & $0.201$ \\ 
+Skeptical & $0.858$ & $0.419$ & $1.755$ & $0.327$ & $0.237$ & $0.197$ \\ 
 \bottomrule

### 2023-02-11 manuscript/manuscript.qmd
starting discussion

@@
 abstract: |
-  A recent randomized clinical trial, INCEPTION, reported patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome.
+  A recent randomized clinical trial, INCEPTION, reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome.
 The current study examines if a Bayesian perspective provides  additional insights.
@@
 The probability of clinical superiority with eCPR ranges from 65.7 - 77.0 % (defined as OR > 1.1).
-A similar analyses using INCEPTION per protocol data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority of approximately 25%.
+A similar analyses using INCEPTION per protocol (PP) data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority not exceeding 25%.
 It is concluded that a Bayesian perspective allows considerable additional insights into the trial analysis and interpretation.
-The substantial probabilities of increased survival, but with considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
+The non-negligible probabilities of increased survival with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
 keywords: 
@@
 and external defibrillation (conventional cardiopulmonary resuscitation
-(CPR)). However, there remains a substantial subset of individuals who
+(cCPR)). However, there remains a substantial subset of individuals who
 do not respond rapidly to these measures and whether more invasive
@@
 to standard advanced cardiac life support (eCPR)) can improve survival and
-diminish anoxic brain injury is a current topic of research. The largest
+diminish anoxic brain injury is a current topic of research. A large
 randomized clinical trial (RCT) examining this question recently
-published their results[@CPR2023a]. For the primary outcome, 30 day
-survival without significant neurological deficit, the authors observed
-an odds ratio of 1.4 (95% confidence interval, 0.5 to 3.5; P = 0.52) in
-favor of extracorporeal CPR for leading to their conclusion "In patients
+published their results[@CPR2023a]. For their primary outcome, 30 day
+survival without significant neurological deficit, an odds ratio of 1.4 (95% confidence interval, 0.5 to 3.5; P = 0.52) in
+favor of extracorporeal CPR was observed leading to the conclusion "In patients
 with refractory out-of-hospital cardiac arrest, extracorporeal CPR and
@@
 Bayesian perspective permits additional insights into the specific
-clinical question regarding any added value of extracorporeal CPR
-following an out-of-hospital arrest in patients refractory to standard
-CPR.
+clinical question regarding any added value of eCPR
+following an out-of-hospital arrest in patients refractory to cCPR.
 
@@
 bias by preserving the prognostic balance afforded by randomization as
-well as assuring the validity of the statistical analyses.
+well as assuring the validity of the accompanying statistical analyses.
 
@@
 These probability statements arise from the posterior distribution
-according to the following equation:
+according to the Bayes Thereom, expressed as follows: 
 $$ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} $$
@@
 Therefore, in addition to the current data summarized by the probability
-of the data (likelihood function) one requires a prior probability
-distribution for each parameter. The mechanics of the Bayesian analyses
-were performed using the Stan programming language [@stan] through the R
-package rstanarm [@rstanarm] and fit a logistic regression model with a
-single treatment parameter, $\theta$. Because our focus is the
-interpretation of the INCEPTION trial alone, our primary analysis used
-rstanarm's default vague parameter priors
+of the data (likelihood function), prior probability
+distributions are required. Because our main focus is the analysis and 
+interpretation of the INCEPTION trial alone, our primary analysis used a default vague parameter prior
 ($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
@@
 analyses that examine the variation in the posterior probability as a
-function of the choice of different prior distributions. Using prior
-information also underscores the important advantage of Bayesian
-analyses to learn sequentially. There were two previous RCTs examining
+function of the choice of different prior distributions. Incorporating prior
+information underscores another important advantage of Bayesian
+analyses, the ability to learn sequentially. There were two previous RCTs examining
 extracorporeal CPR[@RN6759; @RN6751] and while the protocols are not
@@
 
+Therefore, in addition to the vague prior described above, we considered three possible informative priors    
+i) a **combined** prior using all the available prior RCT data[@RN6751;@RN6759]               
+ii) an **enthusiastic** prior, so labelled since this uses only the ARREST data, a trial stopped prematurely for efficacy        
+iii) a **skeptical** prior, so labelled since this uses only the PRAGUE data, a trial  stopped prematurely for futility
+
 This prior
 information of the probability of eCPR success in each previous trial, $X_i$, can be summarized as a normal distribution with a mean equal to the proportion of successes, $\hat{p_i}$ with a standard deviation equal to $$\sqrt{\hat{p_i}*(1-\hat{p_i})}$$
-As  baseline success rates for standard CPR varies markedly between the three studies, it was decided to maintain the INCEPTION control baseline with the vague prior for all analyses and to update only the eCPR arm with prior information.
+As  baseline success rates for cCPR varies markedly between the three studies, it was decided to maintain the INCEPTION control baseline with the vague prior for all analyses and to update only the eCPR arm with prior information.
 
 Posterior distributions are summarized with medians and 95%
-highest-density intervals (credible intervals), defined as the narrowest
+highest-density intervals (credible intervals (CrI)), defined as the narrowest
 interval containing 95% of the probability density function
-[@mcelreath2020]. Bayesian analyses permit not only calculated the posterior probability of
+[@mcelreath2020]. Bayesian analyses permit not only calculations of the posterior probability of
 any additional survival with eCPR (OR \>1.00), but also of clinically meaningful benefits.
-While there is no universal standard for a clinically meaningful benefit, a survival OR \>1.10 may be an acceptable threshold for many.
+While there is no universal definition for a clinically meaningful benefit, a survival OR \>1.10 may be an acceptable threshold for many.
 Bayesian analyses also allows calculation of the probability between any two points.
-For example, rather than simply comparing if the survival of one treatment is better, one can calculate a range of practical equivalence (ROPE).
-While different ranges may be proposed as equivalent, +/- 10% seems a reasonable starting point.
+For example, rather than simply comparing if the survival of one treatment is better than another, one can calculate a range of practical equivalence (ROPE) between treatments.
+While different ranges may be proposed, +/- 10% seems a reasonable small difference that many woul consider as equivalent.
 
@@
 to, regardless of whether or not they dropped out, were fully adhered to
-the treatment or switched to an alternative treatment. ITT analyses can
+the treatment or switched to an alternative treatment. In superiority trials, ITT analyses can
 therefore be seen as a conservative estimate which mirrors clinical
@@
 
-All analyses were executed using `R`[@R]  within the integrated development environment
-of RStudio. Bayesian analyses were performed using the front end `rstanarm` package[@rstanarm] to the `Stan` programming language[stan]. The statistical code can be found on Github
+Posterior distributions were estimated using the Hamiltonian Monte Carlo,a form of 
+Markov Chain Monte Carlo simulations in which the gradient of the log posterior is 
+used to efficiently sample the posterior space.
+This was implemented in `Stan`[@stan] using the front end `rstanarm` package[@rstanarm] by fitting a logistic regression model with a single treatment parameter.
+All analyses were executed using `R`[@R]  within the integrated development environmentof RStudio[@Rstudio].
+The statistical code can be found on Github
 (https://github.com/brophyj/eCPR).
@@
 
-ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6751;@RN6759] that also randomized CPR to eCPR are shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using the default vague priors, produces an odds ratio (OR) 1.32 (95% Credible Interval (CrI) 0.54 - 3.22).
-The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms minimal impact of the vague prior and shows this Bayesian analysis is completely dominated by the observed INCEPTION data.
+ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6759;@RN6751] that also randomized out of hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
+Performing a Bayesian analysis on the INCEPTION trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22.
+The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION data.
 
-One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved survival with eCPR.
-The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,2)`%.
-The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1,2)`%.
+One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved eCPR survival.
+The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%.
+The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1*100,1)`% and the ROPE probability is `r round(vague_eq*100,1)`% (Table 2).
 
@@
 
-Another advantage of a Bayesian approach is the possibility of including previous evidence as informed priors.
-We considered three possible scenarios i) using all the available prior RCT data[@RN6751;@RN6759] to create a combined prior which was expressed as a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`)
-ii) using only the ARREST data, labelled an enthusiastic prior, as this trial was stopped prematurely for efficacy, expressed as a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`)    
-iii) using only the PRAGUE data, labelled a skeptical prior, as this trial was stopped prematurely for futility, expressed as a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`)
+Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively
 
-These different prior probabilities can be updated using the INCEPTION to create posterior distributions that are summarized in Table 2.
-Basically, in all cases the point estimate for the enhanced eCPR survival remains fairly constant but the associated uncertainty is reduced with the additional data as reflected by the narrower 95% CrI.
+These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2.
+The probability for enhanced eCPR survival has increased to 80.4% with a skeptical prior to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.
 The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.
-The graphical presentation of these results is shown in Figure 2.
+The probability that the eCPR survival improvements exceed the proposed 10% clinical threshold are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively.
+The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2).
+Posterior probabilities with the combined prior fell as expected between the results with the skeptical and enthusiastic priors.
+The graphical presentations of these results are shown in Figure 2.
 
-Although INCEPTION did not report a per-protocol analysis, just an analysis may be revealing concerning treatment efficacy, at the risk of an increased risk of bias by not respecting the ITT.
-From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the CPR group compared to 5 survivors from 46 patients receiving eCPR.
-With a vague prior, the probability of increased survival with eCPR is decreased compared to CPR but with wide CrI (OR 0.45, 95% CrI 1.5 - 1.35), limiting any definitive conclusions.
-The probabilities of increased eCPR survival incorporating the previous identified priors is shown in Table 3.
+INCEPTION did not report a per-protocol analysis.
+Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle.
+From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR.
+With a vague prior, the OR of increased survival with eCPR is decreased compared to cCPR but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions.
+The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR.
+The exact probabilities of increased eCPR survival incorporating the previous identified informative priors are shown in Table 3.
 
@@
 ```
+
+# Discussion
+
+This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR.
+The probability that this improvement in survival exceeds 10% is almost 66%.
+The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%.
+Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability.
+In contrast to the INCEPTION original conclusion of similar survival effects between the two treatment, this re-analysis suggests only a modest probability of approximately that survival probabilities between the two techniques are within a reasonably acceptable equivalence range.
+The overall message would appear to be that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit or even less likley clinically meaningful harm has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+
+The INCEPTION researchers are to be congratulated on their trial design, its execution, and addressing an important clinical question in the most challenging of research environments.
+However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge.
+For trails failing to meet statistical significance, those incorrectly labelled "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
 

### 2023-02-11 manuscript/manuscript.tex
starting discussion

@@
 \begin{abstract}
-A recent randomized clinical trial, INCEPTION, reported patients with
+A recent randomized clinical trial, INCEPTION, reported in patients with
 refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and
@@
 \textgreater{} 1.1). A similar analyses using INCEPTION per protocol
-data shows an equivalence probability between 4.7 - 20.2\% with reduced
-probabilities of clinical superiority of approximately 25\%. It is
+(PP) data shows an equivalence probability between 4.7 - 20.2\% with
+reduced probabilities of clinical superiority not exceeding 25\%. It is
 concluded that a Bayesian perspective allows considerable additional
-insights into the trial analysis and interpretation. The substantial
-probabilities of increased survival, but with considerable residual
+insights into the trial analysis and interpretation. The non-negligible
+probabilities of increased survival with their considerable residual
 uncertainties, suggests that additional studies are required before
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 to standard advanced cardiac life support (eCPR)) can improve survival
-and diminish anoxic brain injury is a current topic of research. The
-largest randomized clinical trial (RCT) examining this question recently
-published their results\citep{CPR2023a}. For the primary outcome, 30 day
-survival without significant neurological deficit, the authors observed
-an odds ratio of 1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in
-favor of extracorporeal CPR for leading to their conclusion ``In
-patients with refractory out-of-hospital cardiac arrest, extracorporeal
-CPR and conventional CPR had similar effects on survival with a
-favorable neurological outcome''.\citep{CPR2023a}
+and diminish anoxic brain injury is a current topic of research. A large
+randomized clinical trial (RCT) examining this question recently
+published their results\citep{CPR2023a}. For their primary outcome, 30
+day survival without significant neurological deficit, an odds ratio of
+1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in favor of
+extracorporeal CPR was observed leading to the conclusion ``In patients
+with refractory out-of-hospital cardiac arrest, extracorporeal CPR and
+conventional CPR had similar effects on survival with a favorable
+neurological outcome''.\citep{CPR2023a}
 
@@
 Bayesian perspective permits additional insights into the specific
-clinical question regarding any added value of extracorporeal CPR
-following an out-of-hospital arrest in patients refractory to standard
-CPR.
+clinical question regarding any added value of eCPR following an
+out-of-hospital arrest in patients refractory to cCPR.
 
@@
 minimizing bias by preserving the prognostic balance afforded by
-randomization as well as assuring the validity of the statistical
-analyses.
+randomization as well as assuring the validity of the accompanying
+statistical analyses.
 
@@
 These probability statements arise from the posterior distribution
-according to the following equation:
+according to the Bayes Thereom, expressed as follows:
 \[ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} \]
@@
 Therefore, in addition to the current data summarized by the probability
-of the data (likelihood function) one requires a prior probability
-distribution for each parameter. The mechanics of the Bayesian analyses
-were performed using the Stan programming language \citep{stan} through
-the R package rstanarm \citep{rstanarm} and fit a logistic regression
-model with a single treatment parameter, \(\theta\). Because our focus
-is the interpretation of the INCEPTION trial alone, our primary analysis
-used rstanarm's default vague parameter priors
-(\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring that the
-posterior distribution is dominated by the observed INCEPTION data.
+of the data (likelihood function), prior probability distributions are
+required. Because our main focus is the analysis and interpretation of
+the INCEPTION trial alone, our primary analysis used a default vague
+parameter prior (\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring
+that the posterior distribution is dominated by the observed INCEPTION
+data.
 
@@
 analyses that examine the variation in the posterior probability as a
-function of the choice of different prior distributions. Using prior
-information also underscores the important advantage of Bayesian
-analyses to learn sequentially. There were two previous RCTs examining
-extracorporeal CPR\citep{RN6759, RN6751} and while the protocols are not
-identical, it may be reasonable to allow this data to serve as informed
-priors for the eCPR parameter, which can be updated with the INCEPTION
-data.
+function of the choice of different prior distributions. Incorporating
+prior information underscores another important advantage of Bayesian
+analyses, the ability to learn sequentially. There were two previous
+RCTs examining extracorporeal CPR\citep{RN6759, RN6751} and while the
+protocols are not identical, it may be reasonable to allow this data to
+serve as informed priors for the eCPR parameter, which can be updated
+with the INCEPTION data.
+
+Therefore, in addition to the vague prior described above, we considered
+three possible informative priors\\
+i) a \textbf{combined} prior using all the available prior RCT
+data\citep{RN6751, RN6759}\\
+ii) an \textbf{enthusiastic} prior, so labelled since this uses only the
+ARREST data, a trial stopped prematurely for efficacy\\
+iii) a \textbf{skeptical} prior, so labelled since this uses only the
+PRAGUE data, a trial stopped prematurely for futility
 
@@
 standard deviation equal to \[\sqrt{\hat{p_i}*(1-\hat{p_i})}\] As
-baseline success rates for standard CPR varies markedly between the
-three studies, it was decided to maintain the INCEPTION control baseline
-with the vague prior for all analyses and to update only the eCPR arm
-with prior information.
+baseline success rates for cCPR varies markedly between the three
+studies, it was decided to maintain the INCEPTION control baseline with
+the vague prior for all analyses and to update only the eCPR arm with
+prior information.
 
 Posterior distributions are summarized with medians and 95\%
-highest-density intervals (credible intervals), defined as the narrowest
-interval containing 95\% of the probability density function
-\citep{mcelreath2020}. Bayesian analyses permit not only calculated the
-posterior probability of any additional survival with eCPR (OR
+highest-density intervals (credible intervals (CrI)), defined as the
+narrowest interval containing 95\% of the probability density function
+\citep{mcelreath2020}. Bayesian analyses permit not only calculations of
+the posterior probability of any additional survival with eCPR (OR
 \textgreater1.00), but also of clinically meaningful benefits. While
-there is no universal standard for a clinically meaningful benefit, a
+there is no universal definition for a clinically meaningful benefit, a
 survival OR \textgreater1.10 may be an acceptable threshold for many.
@@
 two points. For example, rather than simply comparing if the survival of
-one treatment is better, one can calculate a range of practical
-equivalence (ROPE). While different ranges may be proposed as
-equivalent, +/- 10\% seems a reasonable starting point.
+one treatment is better than another, one can calculate a range of
+practical equivalence (ROPE) between treatments. While different ranges
+may be proposed, +/- 10\% seems a reasonable small difference that many
+woul consider as equivalent.
 
@@
 were fully adhered to the treatment or switched to an alternative
-treatment. ITT analyses can therefore be seen as a conservative estimate
-which mirrors clinical effectiveness. In contrast, a per protocol (PP)
-analysis accounts for adherence by analyzing only those patients who
-completed the treatment they were originally allocated to. While PP may
-provide additional insights into efficacy, it is subject to bias. Since
-a PP analysis when performed in conjunction with an ITT analysis may
-provide additional insights, this has also been subjected to a Bayesian
-analysis.
+treatment. In superiority trials, ITT analyses can therefore be seen as
+a conservative estimate which mirrors clinical effectiveness. In
+contrast, a per protocol (PP) analysis accounts for adherence by
+analyzing only those patients who completed the treatment they were
+originally allocated to. While PP may provide additional insights into
+efficacy, it is subject to bias. Since a PP analysis when performed in
+conjunction with an ITT analysis may provide additional insights, this
+has also been subjected to a Bayesian analysis.
 
-All analyses were executed using \texttt{R}\citep{R} within the
-integrated development environment of RStudio. Bayesian analyses were
-performed using the front end \texttt{rstanarm} package\citep{rstanarm}
-to the \texttt{Stan} programming language{[}stan{]}. The statistical
-code can be found on Github (https://github.com/brophyj/eCPR).
+Posterior distributions were estimated using the Hamiltonian Monte
+Carlo,a form of Markov Chain Monte Carlo simulations in which the
+gradient of the log posterior is used to efficiently sample the
+posterior space. This was implemented in \texttt{Stan}\citep{stan} using
+the front end \texttt{rstanarm} package\citep{rstanarm} by fitting a
+logistic regression model with a single treatment parameter. All
+analyses were executed using \texttt{R}\citep{R} within the integrated
+development environmentof RStudio\citep{Rstudio}. The statistical code
+can be found on Github (https://github.com/brophyj/eCPR).
 
@@
 ITT data from the INCEPTION trial\citep{CPR2023a} and two other
-pertinent trials\citep{RN6751, RN6759} that also randomized CPR to eCPR
-are shown in Table 1. Performing a Bayesian analysis on the INCEPTION
-trial, using the default vague priors, produces an odds ratio (OR) 1.32
-(95\% Credible Interval (CrI) 0.54 - 3.22). The closeness of this result
-to the original analysis (OR, 1.4; 95\% CI 0.5 - 3.5) confirms minimal
-impact of the vague prior and shows this Bayesian analysis is completely
-dominated by the observed INCEPTION data.
+pertinent trials\citep{RN6759, RN6751} that also randomized out of
+hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
+Performing a Bayesian analysis on the INCEPTION trial, using a default
+vague prior, produces an odds ratio (OR) 1.32 with 95\% CrI 0.54 - 3.22.
+The closeness of this result to the original analysis (OR, 1.4; 95\% CI
+0.5 - 3.5) confirms the minimal impact of the default vague prior and
+reveals a Bayesian analysis completely dominated by the observed
+INCEPTION data.
 
 One of the advantages of a Bayesian approach is the ability to make
-direct probability statements about the estimand of improved survival
-with eCPR. The eCPR probability density function for improved survival
+direct probability statements about the estimand of improved eCPR
+survival. The eCPR probability density function for improved survival
 from INCEPTION data with the default vague prior is displayed in Figure
 1 and reveals that the probability of enhanced survival with eCPR is
-72.72\%. The probability that the improved survival exceeds a 10\%
-improvement is 0.66\%.
+72.7\%. The probability that the improved survival exceeds a 10\%
+improvement is 65.7\% and the ROPE probability is 13.4\% (Table 2).
 
-Another advantage of a Bayesian approach is the possibility of including
-previous evidence as informed priors. We considered three possible
-scenarios i) using all the available prior RCT
-data\citep{RN6751, RN6759} to create a combined prior which was
-expressed as a N(0.32, 0.47)\\
-ii) using only the ARREST data, labelled an enthusiastic prior, as this
-trial was stopped prematurely for efficacy, expressed as a N(0.43,
-0.49)\\
-iii) using only the PRAGUE data, labelled a skeptical prior, as this
-trial was stopped prematurely for futility, expressed as a N(0.31, 0.46)
+Three different informative priors were considered i) a N(0.32, 0.47)
+ii) a N(0.43, 0.49) iii) a N(0.31, 0.46) distributions to represent all
+the combined available RCT data\citep{RN6759, RN6751}, only the
+ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
+respectively
 
-These different prior probabilities can be updated using the INCEPTION
-to create posterior distributions that are summarized in Table 2.
-Basically, in all cases the point estimate for the enhanced eCPR
-survival remains fairly constant but the associated uncertainty is
-reduced with the additional data as reflected by the narrower 95\% CrI.
-The posterior probability of improved eCPR survival goes from 80\% with
-a skeptical prior to 85\% with the enthusiastic prior. The graphical
-presentation of these results is shown in Figure 2.
+These different prior probabilities were updated using with the
+INCEPTION ITT data to create the posterior distributions displayed in
+Table 2. The probability for enhanced eCPR survival has increased to
+80.4\% with a skeptical prior to 84.9\% with the enthusiastic prior and,
+as expected, the associated uncertainty has been reduced, as reflected
+by the narrower 95\% CrI, with the additional data. The posterior
+probability of improved eCPR survival goes from 80\% with a skeptical
+prior to 85\% with the enthusiastic prior. The probability that the eCPR
+survival improvements exceed the proposed 10\% clinical threshold are
+71\% and 77\% for the skeptical and enthusiastic priors, respectively.
+The corresponding ROPE probabilities are 16.8\% and 13.8\% (Table 2).
+Posterior probabilities with the combined prior fell as expected between
+the results with the skeptical and enthusiastic priors. The graphical
+presentations of these results are shown in Figure 2.
 
-Although INCEPTION did not report a per-protocol analysis, just an
-analysis may be revealing concerning treatment efficacy, at the risk of
-an increased risk of bias by not respecting the ITT. From INCEPTION
-Figure S4\citep{CPR2023a}, it appears that the per protocol data for
-mortality is 13 survivors from 61 patients in the CPR group compared to
-5 survivors from 46 patients receiving eCPR. With a vague prior, the
-probability of increased survival with eCPR is decreased compared to CPR
-but with wide CrI (OR 0.45, 95\% CrI 1.5 - 1.35), limiting any
-definitive conclusions. The probabilities of increased eCPR survival
-incorporating the previous identified priors is shown in Table 3.
+INCEPTION did not report a per-protocol analysis. Such an analysis may
+be helpful in assessing treatment efficacy, unfortunately at the risk of
+an increased risk of bias by not respecting the ITT principle. From
+INCEPTION Figure S4\citep{CPR2023a}, it appears that the per protocol
+data for mortality is 13 survivors from 61 patients in the cCPR group
+compared to 5 survivors from 46 patients receiving eCPR. With a vague
+prior, the OR of increased survival with eCPR is decreased compared to
+cCPR but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35), limiting
+any definitive conclusions. The decreased eCPR success rates in the
+INCEPTION PP data results in reduced posterior probabilities of eCPR
+benefit and an increased probability of equivalence or benefit with
+cCPR. The exact probabilities of increased eCPR survival incorporating
+the previous identified informative priors are shown in Table 3.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+This Bayesian analysis of the INCEPTION ITT data alone (with a vague
+prior) suggests the presence of a non-negligible probability of benefit
+with eCPR, in the range of 73\%, compared to cCPR. The probability that
+this improvement in survival exceeds 10\% is almost 66\%. The
+consideration of pertinent prior information from two previous
+RCTs\citep{RN6751, RN6759} in the field results in estimates of improved
+eCPR survival probabilities, with intact neurological status, in the
+vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
+included, the posterior probability of at least a 10\% improved eCPR
+survival is in the range of 71 - 77\% probability. In contrast to the
+INCEPTION original conclusion of similar survival effects between the
+two treatment, this re-analysis suggests only a modest probability of
+approximately that survival probabilities between the two techniques are
+within a reasonably acceptable equivalence range. The overall message
+would appear to be that while any definitive conclusions regarding the
+superiority of eCPR are lacking, the possibility of a clinically
+meaningful benefit or even less likley clinically meaningful harm has
+not been reasonably excluded and continued research is necessary to
+clarifying the remaining uncertainties.
+
+The INCEPTION researchers are to be congratulated on their trial design,
+its execution, and addressing an important clinical question in the most
+challenging of research environments. However the natural constraints of
+standard statistical analyses, also known as null hypothesis
+significance testing, limits the full exploitation of the current data
+and updating of past knowledge. For trails failing to meet statistical
+significance, those incorrectly labelled ``negative'' trials, null
+hypothesis significance testing favors confusion ``between absence of
+evidence and evidence of absence''\citep{RN6765}.
 

### 2023-02-11 manuscript/manuscript.qmd
updated discussion

@@
 Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability.
-In contrast to the INCEPTION original conclusion of similar survival effects between the two treatment, this re-analysis suggests only a modest probability of approximately that survival probabilities between the two techniques are within a reasonably acceptable equivalence range.
-The overall message would appear to be that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit or even less likley clinically meaningful harm has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range.
+Additional uncertainty about the benefits or harms of eCPR arise from the per protocol analysis, both with and without informative priors.
 
@@
 However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge.
-For trails failing to meet statistical significance, those incorrectly labelled "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
+For trials such as INCEPTION that fail to meet statistical significance and are often thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
+Rather than dichotomizing results into statistical significance or not, with its obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
+
+Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to either
+individual data from any of the trials limiting the possibility of examining specific subgroups.
+Also for the per protocol analysis, this data was only available for survival and not the primary outcome that included survival with intact neurological status.
+The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices.
+Given the existence of only 3 trials  differences between the baseline cCPR rates could not be reliably modeled.
+We choose not to perform a traditional Bayesian meta-analysis with
+a semi-informative prior for the between-study variation  because our
+goal was to illustrate how Bayesian principles could be informative
+when applied to the interpretation of new trial data.
+
+The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" is overly simplified and potentially inaccurate.
+This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even somewhat less likely clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
 

### 2023-02-11 manuscript/manuscript.tex
updated discussion

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 INCEPTION original conclusion of similar survival effects between the
-two treatment, this re-analysis suggests only a modest probability of
-approximately that survival probabilities between the two techniques are
-within a reasonably acceptable equivalence range. The overall message
-would appear to be that while any definitive conclusions regarding the
-superiority of eCPR are lacking, the possibility of a clinically
-meaningful benefit or even less likley clinically meaningful harm has
-not been reasonably excluded and continued research is necessary to
-clarifying the remaining uncertainties.
+two treatments, this re-analysis suggests only a modest probability of
+approximately 15\% that survival probabilities between the two
+techniques are within a reasonably acceptable equivalency range.
+Additional uncertainty about the benefits or harms of eCPR arise from
+the per protocol analysis, both with and without informative priors.
 
@@
 significance testing, limits the full exploitation of the current data
-and updating of past knowledge. For trails failing to meet statistical
-significance, those incorrectly labelled ``negative'' trials, null
-hypothesis significance testing favors confusion ``between absence of
-evidence and evidence of absence''\citep{RN6765}.
+and updating of past knowledge. For trials such as INCEPTION that fail
+to meet statistical significance and are often thought of as
+``negative'' trials, null hypothesis significance testing favors
+confusion ``between absence of evidence and evidence of
+absence''\citep{RN6765}. Rather than dichotomizing results into
+statistical significance or not, with its obligatory loss of
+information, Bayesian analysis concentrates on estimation of key outcome
+differences with direct probability measures of their uncertainty.
+
+Although this Bayesian analysis presents the strengths noted above, it
+also had limitations. First, we did not have access to either individual
+data from any of the trials limiting the possibility of examining
+specific subgroups. Also for the per protocol analysis, this data was
+only available for survival and not the primary outcome that included
+survival with intact neurological status. The threshold choices for
+superiority and equivalence may be seen as arbitrary, although the
+analyses can be easily repeated for different choices. Given the
+existence of only 3 trials differences between the baseline cCPR rates
+could not be reliably modeled. We choose not to perform a traditional
+Bayesian meta-analysis with a semi-informative prior for the
+between-study variation because our goal was to illustrate how Bayesian
+principles could be informative when applied to the interpretation of
+new trial data.
+
+The overall message from this study is that standard statistical
+analyses leading to a conclusion of ``similar survival effects'' is
+overly simplified and potentially inaccurate. This Bayesian analysis
+demonstrates clearly that while any definitive conclusions regarding the
+superiority of eCPR are lacking, the possibility of a clinically
+meaningful benefit, or even somewhat less likely clinically meaningful
+harm, has not been reasonably excluded and continued research is
+necessary to clarifying the remaining uncertainties.
 

### 2023-02-12 manuscript/manuscript.qmd
fixed ARREST data 1 of 15 cCPR, added to discussion

@@
 It is concluded that a Bayesian perspective allows considerable additional insights into the trial analysis and interpretation.
-The non-negligible probabilities of increased survival with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
+The non-negligible probabilities of increased survival, or even harm, with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
 keywords: 
@@
 df <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
-             fail = c(52, 56, 15, 8, 108, 86),
-             success = c(10,14, 0, 6,24,38 )) %>% 
+             fail = c(52, 56, 14, 8, 108, 86),
+             success = c(10,14, 1, 6,24,38 )) %>% 
   mutate(total = fail + success,
@@
 The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.
-The probability that the eCPR survival improvements exceed the proposed 10% clinical threshold are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively.
+
+The probability that the eCPR survival improvements exceed a minimum 10% clinical threshold for improvement are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively.
 The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2).
@@
 The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR.
-The exact probabilities of increased eCPR survival incorporating the previous identified informative priors are shown in Table 3.
+The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previous identified informative priors are shown in Table 3.
+
+```{r pred, cache=TRUE, eval=FALSE}
+# some extra based on ROS
+library("bayesplot")
+library("ggplot2")
+library("rstan")
+
+x <- rep(c(0, 1), c(70, 62))  
+y <- rep(c(0, 1, 0, 1), c(14,56,10, 52))  
+simple <- data.frame(x, y)  
+fit <- stan_glm(y ~ x, family=binomial(link="logit"), data=simple) 
+summary(fit)
+
+new <- data.frame(x=c(0, 1))  
+epred <- posterior_epred(fit, newdata=new)  
+diff <- epred[,2] - epred[,1]  
+print(c(mean(diff), sd(diff)))
+
+yrep_1 <- posterior_predict(fit); n_sims <- nrow(yrep_1); subset <- sample(n_sims, 100)  
+bayesplot::ppc_dens_overlay(y = log10(simple$y + 1),
+                            yrep = log10(yrep_1[subset, ] + 1)) +
+  theme(axis.line.y = element_blank(),text = element_text(family = "sans")) +
+  labs(title = "Binomial model",x = "log10(y + 1)")
+
+### other functions
+pp_check (fit1,  nreps=30)
+plot (fit1, "dens")
+
+```
 
@@
 
-The INCEPTION researchers are to be congratulated on their trial design, its execution, and addressing an important clinical question in the most challenging of research environments.
-However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge.
+The INCEPTION researchers have ddressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion.
+However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data and prevents a full and comprehensive exploitation and updating of past knowledge.
 For trials such as INCEPTION that fail to meet statistical significance and are often thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
@@
 The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" is overly simplified and potentially inaccurate.
-This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even somewhat less likely clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+The final dilemma confronting clinicians is that even a Bayesian analysis of a randomized trial provides only probability estimates of an average treatment effect and not the elusive individual treatment effect.
 
@@
        caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 73.2%") +
-  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 81.2% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
+  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 81.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
   geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .86)) +

### 2023-02-12 manuscript/manuscript.tex
fixed ARREST data 1 of 15 cCPR, added to discussion

@@
 insights into the trial analysis and interpretation. The non-negligible
-probabilities of increased survival with their considerable residual
-uncertainties, suggests that additional studies are required before
-concluding that eCPR and cCPR have similar survival effects.
+probabilities of increased survival, or even harm, with their
+considerable residual uncertainties, suggests that additional studies
+are required before concluding that eCPR and cCPR have similar survival
+effects.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 probability of improved eCPR survival goes from 80\% with a skeptical
-prior to 85\% with the enthusiastic prior. The probability that the eCPR
-survival improvements exceed the proposed 10\% clinical threshold are
-71\% and 77\% for the skeptical and enthusiastic priors, respectively.
-The corresponding ROPE probabilities are 16.8\% and 13.8\% (Table 2).
-Posterior probabilities with the combined prior fell as expected between
-the results with the skeptical and enthusiastic priors. The graphical
-presentations of these results are shown in Figure 2.
+prior to 85\% with the enthusiastic prior.
+
+The probability that the eCPR survival improvements exceed a minimum
+10\% clinical threshold for improvement are 71\% and 77\% for the
+skeptical and enthusiastic priors, respectively. The corresponding ROPE
+probabilities are 16.8\% and 13.8\% (Table 2). Posterior probabilities
+with the combined prior fell as expected between the results with the
+skeptical and enthusiastic priors. The graphical presentations of these
+results are shown in Figure 2.
 
@@
 benefit and an increased probability of equivalence or benefit with
-cCPR. The exact probabilities of increased eCPR survival incorporating
-the previous identified informative priors are shown in Table 3.
+cCPR. The probabilities of eCPR survival compared to cCPR using the per
+protocol data and incorporating the previous identified informative
+priors are shown in Table 3.
 
@@
 
-The INCEPTION researchers are to be congratulated on their trial design,
-its execution, and addressing an important clinical question in the most
-challenging of research environments. However the natural constraints of
-standard statistical analyses, also known as null hypothesis
-significance testing, limits the full exploitation of the current data
-and updating of past knowledge. For trials such as INCEPTION that fail
-to meet statistical significance and are often thought of as
-``negative'' trials, null hypothesis significance testing favors
-confusion ``between absence of evidence and evidence of
+The INCEPTION researchers have ddressed an important clinical question
+in the most challenging of research environments and are to be
+congratulated on their trial design, its execution, and a nuanced
+discussion. However the natural constraints of standard statistical
+analyses, also known as null hypothesis significance testing, limits the
+quantitative appreciation of their data and prevents a full and
+comprehensive exploitation and updating of past knowledge. For trials
+such as INCEPTION that fail to meet statistical significance and are
+often thought of as ``negative'' trials, null hypothesis significance
+testing favors confusion ``between absence of evidence and evidence of
 absence''\citep{RN6765}. Rather than dichotomizing results into
@@
 superiority of eCPR are lacking, the possibility of a clinically
-meaningful benefit, or even somewhat less likely clinically meaningful
-harm, has not been reasonably excluded and continued research is
-necessary to clarifying the remaining uncertainties.
+meaningful benefit, or even less likely the possibility of clinically
+meaningful harm, has not been reasonably excluded and continued research
+is necessary to clarifying the remaining uncertainties. The final
+dilemma confronting clinicians is that even a Bayesian analysis of a
+randomized trial provides only probability estimates of an average
+treatment effect and not the elusive individual treatment effect.
 
@@
 INCEPTION & 52 & 56 & 10 & 14 \\ 
-ARREST & 15 & 8 & 0 & 6 \\ 
+ARREST & 14 & 8 & 1 & 6 \\ 
 PRAGUE & 108 & 86 & 24 & 38 \\ 

### 2023-02-12 manuscript/manuscript.qmd
convergence added

@@
   A recent randomized clinical trial, INCEPTION, reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome.
-The current study examines if a Bayesian perspective provides  additional insights.
+The current study examines if a Bayesian perspective provides  additional quantative insights.
 Depending on the prior selected, the Bayesian approach for the INCEPTION intention-to treat (ITT) analysis shows an equivalence probability between 13.4 - 16.8% (defined as 1 / 1.1 < odds ratio (OR) < 1.1).
@@
 A similar analyses using INCEPTION per protocol (PP) data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority not exceeding 25%.
-It is concluded that a Bayesian perspective allows considerable additional insights into the trial analysis and interpretation.
-The non-negligible probabilities of increased survival, or even harm, with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
+It is concluded that a Bayesian perspective allows considerable additional quantative insights into the trial analysis and interpretation.
+The non-negligible probabilities of increased survival, or even harm, with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar average survival effects.
 keywords: 
@@
 All analyses were executed using `R`[@R]  within the integrated development environmentof RStudio[@Rstudio].
+Model convergence was assessed by examination of the Monte Carlo Standard Error < 10% of the posterior standard deviation, $n_{eff}$ an estimate of the effective number of independent draws from the posterior distribution of the estimand > 10% maximum and $\hat{R}$  a measures the ratio of the average variance of samples within each chain to the variance of the pooled samples across chains < 1.1.
 The statistical code can be found on Github
@@
 fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
-fit1_sum <- summary(fit1, digits=3)
+fit1_sum <- summary(fit1, digits=3) # check convergence
 conf_95 <- c(exp(fit1_sum[2]),exp(fit1_sum[2]+c(-1,1)*1.96*fit1_sum[10]))
@@
 # help('prior_summary.stanreg') 
+# plot(fit1)
 # P > 1.1
@@
 The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices.
-Given the existence of only 3 trials  differences between the baseline cCPR rates could not be reliably modeled.
+Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled.
 We choose not to perform a traditional Bayesian meta-analysis with
@@
 goal was to illustrate how Bayesian principles could be informative
-when applied to the interpretation of new trial data.
+when applied to the analysis and interpretation of new trial data, both in the situation of no and considerable previous knowledge.
 
-The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" is overly simplified and potentially inaccurate.
+The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate.
 This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.

### 2023-02-12 manuscript/manuscript.tex
convergence added

@@
 neurologic outcome. The current study examines if a Bayesian perspective
-provides additional insights. Depending on the prior selected, the
-Bayesian approach for the INCEPTION intention-to treat (ITT) analysis
-shows an equivalence probability between 13.4 - 16.8\% (defined as 1 /
-1.1 \textless{} odds ratio (OR) \textless{} 1.1). The probability of
-clinical superiority with eCPR ranges from 65.7 - 77.0 \% (defined as OR
-\textgreater{} 1.1). A similar analyses using INCEPTION per protocol
-(PP) data shows an equivalence probability between 4.7 - 20.2\% with
-reduced probabilities of clinical superiority not exceeding 25\%. It is
-concluded that a Bayesian perspective allows considerable additional
-insights into the trial analysis and interpretation. The non-negligible
-probabilities of increased survival, or even harm, with their
-considerable residual uncertainties, suggests that additional studies
-are required before concluding that eCPR and cCPR have similar survival
-effects.
+provides additional quantative insights. Depending on the prior
+selected, the Bayesian approach for the INCEPTION intention-to treat
+(ITT) analysis shows an equivalence probability between 13.4 - 16.8\%
+(defined as 1 / 1.1 \textless{} odds ratio (OR) \textless{} 1.1). The
+probability of clinical superiority with eCPR ranges from 65.7 - 77.0 \%
+(defined as OR \textgreater{} 1.1). A similar analyses using INCEPTION
+per protocol (PP) data shows an equivalence probability between 4.7 -
+20.2\% with reduced probabilities of clinical superiority not exceeding
+25\%. It is concluded that a Bayesian perspective allows considerable
+additional quantative insights into the trial analysis and
+interpretation. The non-negligible probabilities of increased survival,
+or even harm, with their considerable residual uncertainties, suggests
+that additional studies are required before concluding that eCPR and
+cCPR have similar average survival effects.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, frame hidden, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 analyses were executed using \texttt{R}\citep{R} within the integrated
-development environmentof RStudio\citep{Rstudio}. The statistical code
-can be found on Github (https://github.com/brophyj/eCPR).
+development environmentof RStudio\citep{Rstudio}. Model convergence was
+assessed by examination of the Monte Carlo Standard Error \textless{}
+10\% of the posterior standard deviation, \(n_{eff}\) an estimate of the
+effective number of independent draws from the posterior distribution of
+the estimand \textgreater{} 10\% maximum and \(\hat{R}\) a measures the
+ratio of the average variance of samples within each chain to the
+variance of the pooled samples across chains \textless{} 1.1. The
+statistical code can be found on Github
+(https://github.com/brophyj/eCPR).
 
@@
 analyses can be easily repeated for different choices. Given the
-existence of only 3 trials differences between the baseline cCPR rates
-could not be reliably modeled. We choose not to perform a traditional
-Bayesian meta-analysis with a semi-informative prior for the
+existence of only 3 trials estimating the very divergent baseline cCPR
+rates could not be reliably modeled. We choose not to perform a
+traditional Bayesian meta-analysis with a semi-informative prior for the
 between-study variation because our goal was to illustrate how Bayesian
-principles could be informative when applied to the interpretation of
-new trial data.
+principles could be informative when applied to the analysis and
+interpretation of new trial data, both in the situation of no and
+considerable previous knowledge.
 
 The overall message from this study is that standard statistical
-analyses leading to a conclusion of ``similar survival effects'' is
+analyses leading to a conclusion of ``similar survival effects'' may be
 overly simplified and potentially inaccurate. This Bayesian analysis

### 2023-02-12 manuscript/manuscript.qmd
bayesian guideline reporting

@@
 Model convergence was assessed by examination of the Monte Carlo Standard Error < 10% of the posterior standard deviation, $n_{eff}$ an estimate of the effective number of independent draws from the posterior distribution of the estimand > 10% maximum and $\hat{R}$  a measures the ratio of the average variance of samples within each chain to the variance of the pooled samples across chains < 1.1.
+Reporting has followed the Bayesian Analysis Reporting Guidelines[@RN6637].
 The statistical code can be found on Github
@@
 # help('prior_summary.stanreg') 
+# mcmc_pairs(as.array(fit1),  pars = c("(Intercept)",  "TxeCPR" ),off_diag_args = list(size = 0.75))
+# mcmc_trace(fit1)
 # plot(fit1)

### 2023-02-12 manuscript/manuscript.tex
bayesian guideline reporting

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, frame hidden, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 ratio of the average variance of samples within each chain to the
-variance of the pooled samples across chains \textless{} 1.1. The
-statistical code can be found on Github
+variance of the pooled samples across chains \textless{} 1.1. Reporting
+has followed the Bayesian Analysis Reporting Guidelines\citep{RN6637}.
+The statistical code can be found on Github
 (https://github.com/brophyj/eCPR).

### 2023-02-12 manuscript/manuscript.qmd
getting closer

@@
 
-The INCEPTION researchers have ddressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion.
-However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data and prevents a full and comprehensive exploitation and updating of past knowledge.
-For trials such as INCEPTION that fail to meet statistical significance and are often thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
-Rather than dichotomizing results into statistical significance or not, with its obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
+The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion.
+However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge.
+For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
+Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
 
-Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to either
+Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
 individual data from any of the trials limiting the possibility of examining specific subgroups.
-Also for the per protocol analysis, this data was only available for survival and not the primary outcome that included survival with intact neurological status.
+Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status.
 The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices.
@@
 goal was to illustrate how Bayesian principles could be informative
-when applied to the analysis and interpretation of new trial data, both in the situation of no and considerable previous knowledge.
+when applied to the analysis and interpretation of new trial data, both in the situations of no and considerable previous knowledge.
 
-The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate.
-This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
-The final dilemma confronting clinicians is that even a Bayesian analysis of a randomized trial provides only probability estimates of an average treatment effect and not the elusive individual treatment effect.
+The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate.
+This Bayesian analysis demonstrates that while any definitive conclusions regarding the superiority, inferiority or equivalence of eCPR to cCPR are lacking.
+Rather the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for an average treatment effect and not for the elusive individual treatment effect.
 

### 2023-02-12 manuscript/manuscript.tex
getting closer

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-The INCEPTION researchers have ddressed an important clinical question
+The INCEPTION researchers have addressed an important clinical question
 in the most challenging of research environments and are to be
@@
 analyses, also known as null hypothesis significance testing, limits the
-quantitative appreciation of their data and prevents a full and
+quantitative appreciation of their data, and prevents a full and
 comprehensive exploitation and updating of past knowledge. For trials
 such as INCEPTION that fail to meet statistical significance and are
-often thought of as ``negative'' trials, null hypothesis significance
-testing favors confusion ``between absence of evidence and evidence of
-absence''\citep{RN6765}. Rather than dichotomizing results into
-statistical significance or not, with its obligatory loss of
+often incorrectly thought of as ``negative'' trials, null hypothesis
+significance testing favors confusion ``between absence of evidence and
+evidence of absence''\citep{RN6765}. Rather than dichotomizing results
+into statistical significance or not, with an obligatory loss of
 information, Bayesian analysis concentrates on estimation of key outcome
@@
 Although this Bayesian analysis presents the strengths noted above, it
-also had limitations. First, we did not have access to either individual
-data from any of the trials limiting the possibility of examining
-specific subgroups. Also for the per protocol analysis, this data was
-only available for survival and not the primary outcome that included
-survival with intact neurological status. The threshold choices for
-superiority and equivalence may be seen as arbitrary, although the
+also had limitations. First, we did not have access to individual data
+from any of the trials limiting the possibility of examining specific
+subgroups. Also for the per protocol analysis, this data was only
+available for survival and not the primary INCEPTION outcome that
+included survival with intact neurological status. The threshold choices
+for superiority and equivalence may be seen as arbitrary, although the
 analyses can be easily repeated for different choices. Given the
@@
 principles could be informative when applied to the analysis and
-interpretation of new trial data, both in the situation of no and
+interpretation of new trial data, both in the situations of no and
 considerable previous knowledge.
 
-The overall message from this study is that standard statistical
-analyses leading to a conclusion of ``similar survival effects'' may be
-overly simplified and potentially inaccurate. This Bayesian analysis
-demonstrates clearly that while any definitive conclusions regarding the
-superiority of eCPR are lacking, the possibility of a clinically
-meaningful benefit, or even less likely the possibility of clinically
-meaningful harm, has not been reasonably excluded and continued research
-is necessary to clarifying the remaining uncertainties. The final
-dilemma confronting clinicians is that even a Bayesian analysis of a
-randomized trial provides only probability estimates of an average
-treatment effect and not the elusive individual treatment effect.
+The ``take home'' message from this study is that standard statistical
+analyses resulting in a conclusion of ``similar survival effects'' may
+be overly simplified and potentially inaccurate. This Bayesian analysis
+demonstrates that while any definitive conclusions regarding the
+superiority, inferiority or equivalence of eCPR to cCPR are lacking.
+Rather the possibility of a clinically meaningful benefit, or even less
+likely the possibility of clinically meaningful harm, has not been
+reasonably excluded and continued research is necessary to clarifying
+the remaining uncertainties. The final dilemma confronting clinicians is
+that even Bayesian analyses of randomized trials provide only
+probability estimates for an average treatment effect and not for the
+elusive individual treatment effect.
 

### 2023-02-12 manuscript/manuscript.qmd
minor touch ups

@@
 The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate.
-This Bayesian analysis demonstrates that while any definitive conclusions regarding the superiority, inferiority or equivalence of eCPR to cCPR are lacking.
-Rather the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
-The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for an average treatment effect and not for the elusive individual treatment effect.
+This Bayesian analysis demonstrates that any definitive conclusions regarding the superiority, inferiority, or equivalence of eCPR to cCPR are lacking.
+Rather the possibility of a clinically meaningful benefit, or less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
+The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for *average* treatment effect and not for the elusive *individual* treatment effect.
 

### 2023-02-12 manuscript/manuscript.tex
minor touch ups

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 be overly simplified and potentially inaccurate. This Bayesian analysis
-demonstrates that while any definitive conclusions regarding the
-superiority, inferiority or equivalence of eCPR to cCPR are lacking.
-Rather the possibility of a clinically meaningful benefit, or even less
-likely the possibility of clinically meaningful harm, has not been
-reasonably excluded and continued research is necessary to clarifying
-the remaining uncertainties. The final dilemma confronting clinicians is
-that even Bayesian analyses of randomized trials provide only
-probability estimates for an average treatment effect and not for the
-elusive individual treatment effect.
+demonstrates that any definitive conclusions regarding the superiority,
+inferiority, or equivalence of eCPR to cCPR are lacking. Rather the
+possibility of a clinically meaningful benefit, or less likely the
+possibility of clinically meaningful harm, has not been reasonably
+excluded and continued research is necessary to clarifying the remaining
+uncertainties. The final dilemma confronting clinicians is that even
+Bayesian analyses of randomized trials provide only probability
+estimates for \emph{average} treatment effect and not for the elusive
+\emph{individual} treatment effect.
 

### 2023-02-13 manuscript/manuscript.qmd
grammer corrections

@@
 
-Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively
-
+Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively.
 These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2.
-The probability for enhanced eCPR survival has increased to 80.4% with a skeptical prior to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.
-The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior.
+The posteriorprobability for enhanced eCPR survival has increased to 80.4% with a skeptical prior, to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.
 
@@
 The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2).
-Posterior probabilities with the combined prior fell as expected between the results with the skeptical and enthusiastic priors.
+Posterior probabilities with the combined prior were as expected between the results with the skeptical and enthusiastic priors.
 The graphical presentations of these results are shown in Figure 2.
@@
 From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR.
-With a vague prior, the OR of increased survival with eCPR is decreased compared to cCPR but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions.
+With a vague prior, the OR of increased survival with eCPR compared to cCPR is decreased but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions.
 The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR.
-The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previous identified informative priors are shown in Table 3.
+The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previously identified informative priors are shown in Table 3.
 
@@
 
-This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR.
+This Bayesian analysis of the INCEPTION ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR.
 The probability that this improvement in survival exceeds 10% is almost 66%.
-The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%.
+The consideration of pertinent prior information from two previous comparable RCTs[@RN6751;@RN6759] results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%.
 Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability.
 In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range.
-Additional uncertainty about the benefits or harms of eCPR arise from the per protocol analysis, both with and without informative priors.
+Additional uncertainty about the benefits or harms of eCPR do arise with the per protocol analysis, both with and without informative priors.
 
@@
 For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
-Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
+Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information and understanding, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
 
@@
 goal was to illustrate how Bayesian principles could be informative
-when applied to the analysis and interpretation of new trial data, both in the situations of no and considerable previous knowledge.
+when applied to the analysis and interpretation of new trial data, in situations both with and without previous knowledge.
 
-The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate.
-This Bayesian analysis demonstrates that any definitive conclusions regarding the superiority, inferiority, or equivalence of eCPR to cCPR are lacking.
-Rather the possibility of a clinically meaningful benefit, or less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.
-The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for *average* treatment effect and not for the elusive *individual* treatment effect.
+The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects of eCPR to cCPR" may be overly simplified and potentially inaccurate.
+This Bayesian analysis demonstrates that at present definitive conclusions regarding the superiority, inferiority, or equivalence are impossible.
+Rather the possibility of a clinically meaningful benefit, or less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the residual uncertainties.
+The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for *average* treatment effects and not for the elusive *individual* treatment effect.
 

### 2023-02-13 manuscript/manuscript.tex
grammer corrections

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 ARREST\citep{RN6751} data and only the PRAGUE\citep{RN6759} data,
-respectively
-
-These different prior probabilities were updated using with the
-INCEPTION ITT data to create the posterior distributions displayed in
-Table 2. The probability for enhanced eCPR survival has increased to
-80.4\% with a skeptical prior to 84.9\% with the enthusiastic prior and,
-as expected, the associated uncertainty has been reduced, as reflected
-by the narrower 95\% CrI, with the additional data. The posterior
-probability of improved eCPR survival goes from 80\% with a skeptical
-prior to 85\% with the enthusiastic prior.
+respectively. These different prior probabilities were updated using
+with the INCEPTION ITT data to create the posterior distributions
+displayed in Table 2. The posteriorprobability for enhanced eCPR
+survival has increased to 80.4\% with a skeptical prior, to 84.9\% with
+the enthusiastic prior and, as expected, the associated uncertainty has
+been reduced, as reflected by the narrower 95\% CrI, with the additional
+data.
 
@@
 probabilities are 16.8\% and 13.8\% (Table 2). Posterior probabilities
-with the combined prior fell as expected between the results with the
+with the combined prior were as expected between the results with the
 skeptical and enthusiastic priors. The graphical presentations of these
@@
 compared to 5 survivors from 46 patients receiving eCPR. With a vague
-prior, the OR of increased survival with eCPR is decreased compared to
-cCPR but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35), limiting
-any definitive conclusions. The decreased eCPR success rates in the
-INCEPTION PP data results in reduced posterior probabilities of eCPR
+prior, the OR of increased survival with eCPR compared to cCPR is
+decreased but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35),
+limiting any definitive conclusions. The decreased eCPR success rates in
+the INCEPTION PP data results in reduced posterior probabilities of eCPR
 benefit and an increased probability of equivalence or benefit with
 cCPR. The probabilities of eCPR survival compared to cCPR using the per
-protocol data and incorporating the previous identified informative
+protocol data and incorporating the previously identified informative
 priors are shown in Table 3.
@@
 
-This Bayesian analysis of the INCEPTION ITT data alone (with a vague
-prior) suggests the presence of a non-negligible probability of benefit
-with eCPR, in the range of 73\%, compared to cCPR. The probability that
-this improvement in survival exceeds 10\% is almost 66\%. The
-consideration of pertinent prior information from two previous
-RCTs\citep{RN6751, RN6759} in the field results in estimates of improved
+This Bayesian analysis of the INCEPTION ITT data alone (analysis with a
+vague prior) suggests the presence of a non-negligible mortality
+probability benefit with eCPR, in the range of 73\%, compared to cCPR.
+The probability that this improvement in survival exceeds 10\% is almost
+66\%. The consideration of pertinent prior information from two previous
+comparable RCTs\citep{RN6751, RN6759} results in estimates of improved
 eCPR survival probabilities, with intact neurological status, in the
@@
 techniques are within a reasonably acceptable equivalency range.
-Additional uncertainty about the benefits or harms of eCPR arise from
+Additional uncertainty about the benefits or harms of eCPR do arise with
 the per protocol analysis, both with and without informative priors.
@@
 into statistical significance or not, with an obligatory loss of
-information, Bayesian analysis concentrates on estimation of key outcome
-differences with direct probability measures of their uncertainty.
+information and understanding, Bayesian analysis concentrates on
+estimation of key outcome differences with direct probability measures
+of their uncertainty.
 
@@
 principles could be informative when applied to the analysis and
-interpretation of new trial data, both in the situations of no and
-considerable previous knowledge.
+interpretation of new trial data, in situations both with and without
+previous knowledge.
 
 The ``take home'' message from this study is that standard statistical
-analyses resulting in a conclusion of ``similar survival effects'' may
-be overly simplified and potentially inaccurate. This Bayesian analysis
-demonstrates that any definitive conclusions regarding the superiority,
-inferiority, or equivalence of eCPR to cCPR are lacking. Rather the
-possibility of a clinically meaningful benefit, or less likely the
-possibility of clinically meaningful harm, has not been reasonably
-excluded and continued research is necessary to clarifying the remaining
-uncertainties. The final dilemma confronting clinicians is that even
-Bayesian analyses of randomized trials provide only probability
-estimates for \emph{average} treatment effect and not for the elusive
-\emph{individual} treatment effect.
+analyses resulting in a conclusion of ``similar survival effects of eCPR
+to cCPR'' may be overly simplified and potentially inaccurate. This
+Bayesian analysis demonstrates that at present definitive conclusions
+regarding the superiority, inferiority, or equivalence are impossible.
+Rather the possibility of a clinically meaningful benefit, or less
+likely the possibility of clinically meaningful harm, has not been
+reasonably excluded and continued research is necessary to clarifying
+the residual uncertainties. The final dilemma confronting clinicians is
+that even Bayesian analyses of randomized trials provide only
+probability estimates for \emph{average} treatment effects and not for
+the elusive \emph{individual} treatment effect.
 

### 2023-02-13 manuscript/manuscript.qmd
file for Circulation

@@
     journal:
-      name: To be determined
+      name: Circulation
       formatting: preprint
@@
 distributions are required. Because our main focus is the analysis and 
-interpretation of the INCEPTION trial alone, our primary analysis used a default vague parameter prior
+interpretation of the INCEPTION trial[@CPR2023a] alone, our primary analysis used a default vague parameter prior
 ($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
-posterior distribution is dominated by the observed INCEPTION data.
+posterior distribution is dominated by the observed INCEPTION[@CPR2023a] data.
 
@@
 identical, it may be reasonable to allow this data to serve as 
-informed priors for the eCPR parameter, which can be updated with the INCEPTION data.
+informed priors for the eCPR parameter, which can be updated with the INCEPTION[@CPR2023a] data.
 
@@
 ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6759;@RN6751] that also randomized out of hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22.
-The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION data.
+Performing a Bayesian analysis on the INCEPTION[@CPR2023a] trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22.
+The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION[@CPR2023a] data.
 
 One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved eCPR survival.
-The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%.
+The eCPR probability density function for improved survival from INCEPTION[@CPR2023a] data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%.
 The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1*100,1)`% and the ROPE probability is `r round(vague_eq*100,1)`% (Table 2).
@@
 Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively.
-These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2.
+These different prior probabilities were updated using with the INCEPTION[@CPR2023a] ITT data to create the posterior distributions displayed in Table 2.
 The posteriorprobability for enhanced eCPR survival has increased to 80.4% with a skeptical prior, to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data.
@@
 
-INCEPTION did not report a per-protocol analysis.
+INCEPTION[@CPR2023a] did not report a per-protocol analysis.
 Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle.
-From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR.
+From INCEPTION[@CPR2023a] Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR.
 With a vague prior, the OR of increased survival with eCPR compared to cCPR is decreased but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions.
-The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR.
+The decreased eCPR success rates in the INCEPTION[@CPR2023a] per protocol data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR.
 The probabilities of eCPR survival compared to cCPR using the per protocol data and incorporating the previously identified informative priors are shown in Table 3.
@@
 
-This Bayesian analysis of the INCEPTION ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR.
+This Bayesian analysis of the INCEPTION[@CPR2023a] ITT data alone (analysis with a vague prior) suggests the presence of a non-negligible mortality probability benefit with eCPR, in the range of 73%, compared to cCPR.
 The probability that this improvement in survival exceeds 10% is almost 66%.
@@
 Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability.
-In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range.
+In contrast to the INCEPTION[@CPR2023a] original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range.
 Additional uncertainty about the benefits or harms of eCPR do arise with the per protocol analysis, both with and without informative priors.
 
-The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion.
+The INCEPTION[@CPR2023a] researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion.
 However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge.
-For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
+For trials such as INCEPTION[@CPR2023a] that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].
 Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information and understanding, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.
@@
 individual data from any of the trials limiting the possibility of examining specific subgroups.
-Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status.
+Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION[@CPR2023a] outcome that included survival with intact neurological status.
 The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices.

### 2023-02-13 manuscript/manuscript.tex
file for Circulation

@@
 \makeatother
-\journal{To be determined}
+\journal{Circulation}
 \ifLuaTeX
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 required. Because our main focus is the analysis and interpretation of
-the INCEPTION trial alone, our primary analysis used a default vague
-parameter prior (\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring
-that the posterior distribution is dominated by the observed INCEPTION
-data.
+the INCEPTION trial\citep{CPR2023a} alone, our primary analysis used a
+default vague parameter prior (\(log(\theta) \sim Normal [0, 2.50]\),
+thereby assuring that the posterior distribution is dominated by the
+observed INCEPTION\citep{CPR2023a} data.
 
@@
 serve as informed priors for the eCPR parameter, which can be updated
-with the INCEPTION data.
+with the INCEPTION\citep{CPR2023a} data.
 
@@
 hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1.
-Performing a Bayesian analysis on the INCEPTION trial, using a default
-vague prior, produces an odds ratio (OR) 1.32 with 95\% CrI 0.54 - 3.22.
-The closeness of this result to the original analysis (OR, 1.4; 95\% CI
-0.5 - 3.5) confirms the minimal impact of the default vague prior and
-reveals a Bayesian analysis completely dominated by the observed
-INCEPTION data.
+Performing a Bayesian analysis on the INCEPTION\citep{CPR2023a} trial,
+using a default vague prior, produces an odds ratio (OR) 1.32 with 95\%
+CrI 0.54 - 3.22. The closeness of this result to the original analysis
+(OR, 1.4; 95\% CI 0.5 - 3.5) confirms the minimal impact of the default
+vague prior and reveals a Bayesian analysis completely dominated by the
+observed INCEPTION\citep{CPR2023a} data.
 
@@
 survival. The eCPR probability density function for improved survival
-from INCEPTION data with the default vague prior is displayed in Figure
-1 and reveals that the probability of enhanced survival with eCPR is
-72.7\%. The probability that the improved survival exceeds a 10\%
-improvement is 65.7\% and the ROPE probability is 13.4\% (Table 2).
+from INCEPTION\citep{CPR2023a} data with the default vague prior is
+displayed in Figure 1 and reveals that the probability of enhanced
+survival with eCPR is 72.7\%. The probability that the improved survival
+exceeds a 10\% improvement is 65.7\% and the ROPE probability is 13.4\%
+(Table 2).
 
@@
 respectively. These different prior probabilities were updated using
-with the INCEPTION ITT data to create the posterior distributions
-displayed in Table 2. The posteriorprobability for enhanced eCPR
-survival has increased to 80.4\% with a skeptical prior, to 84.9\% with
-the enthusiastic prior and, as expected, the associated uncertainty has
-been reduced, as reflected by the narrower 95\% CrI, with the additional
-data.
+with the INCEPTION\citep{CPR2023a} ITT data to create the posterior
+distributions displayed in Table 2. The posteriorprobability for
+enhanced eCPR survival has increased to 80.4\% with a skeptical prior,
+to 84.9\% with the enthusiastic prior and, as expected, the associated
+uncertainty has been reduced, as reflected by the narrower 95\% CrI,
+with the additional data.
 
@@
 
-INCEPTION did not report a per-protocol analysis. Such an analysis may
-be helpful in assessing treatment efficacy, unfortunately at the risk of
-an increased risk of bias by not respecting the ITT principle. From
-INCEPTION Figure S4\citep{CPR2023a}, it appears that the per protocol
-data for mortality is 13 survivors from 61 patients in the cCPR group
-compared to 5 survivors from 46 patients receiving eCPR. With a vague
-prior, the OR of increased survival with eCPR compared to cCPR is
-decreased but with very wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35),
-limiting any definitive conclusions. The decreased eCPR success rates in
-the INCEPTION PP data results in reduced posterior probabilities of eCPR
-benefit and an increased probability of equivalence or benefit with
-cCPR. The probabilities of eCPR survival compared to cCPR using the per
-protocol data and incorporating the previously identified informative
-priors are shown in Table 3.
+INCEPTION\citep{CPR2023a} did not report a per-protocol analysis. Such
+an analysis may be helpful in assessing treatment efficacy,
+unfortunately at the risk of an increased risk of bias by not respecting
+the ITT principle. From INCEPTION\citep{CPR2023a} Figure
+S4\citep{CPR2023a}, it appears that the per protocol data for mortality
+is 13 survivors from 61 patients in the cCPR group compared to 5
+survivors from 46 patients receiving eCPR. With a vague prior, the OR of
+increased survival with eCPR compared to cCPR is decreased but with very
+wide CrI (OR 0.45, 95\% CrI 0.15 - 1.35), limiting any definitive
+conclusions. The decreased eCPR success rates in the
+INCEPTION\citep{CPR2023a} per protocol data results in reduced posterior
+probabilities of eCPR benefit and an increased probability of
+equivalence or benefit with cCPR. The probabilities of eCPR survival
+compared to cCPR using the per protocol data and incorporating the
+previously identified informative priors are shown in Table 3.
 
@@
 
-This Bayesian analysis of the INCEPTION ITT data alone (analysis with a
-vague prior) suggests the presence of a non-negligible mortality
-probability benefit with eCPR, in the range of 73\%, compared to cCPR.
-The probability that this improvement in survival exceeds 10\% is almost
-66\%. The consideration of pertinent prior information from two previous
-comparable RCTs\citep{RN6751, RN6759} results in estimates of improved
-eCPR survival probabilities, with intact neurological status, in the
-vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
-included, the posterior probability of at least a 10\% improved eCPR
-survival is in the range of 71 - 77\% probability. In contrast to the
-INCEPTION original conclusion of similar survival effects between the
-two treatments, this re-analysis suggests only a modest probability of
-approximately 15\% that survival probabilities between the two
-techniques are within a reasonably acceptable equivalency range.
-Additional uncertainty about the benefits or harms of eCPR do arise with
-the per protocol analysis, both with and without informative priors.
+This Bayesian analysis of the INCEPTION\citep{CPR2023a} ITT data alone
+(analysis with a vague prior) suggests the presence of a non-negligible
+mortality probability benefit with eCPR, in the range of 73\%, compared
+to cCPR. The probability that this improvement in survival exceeds 10\%
+is almost 66\%. The consideration of pertinent prior information from
+two previous comparable RCTs\citep{RN6751, RN6759} results in estimates
+of improved eCPR survival probabilities, with intact neurological
+status, in the vicinity of 80-85\%. Moreover depending on the choice of
+prior beliefs included, the posterior probability of at least a 10\%
+improved eCPR survival is in the range of 71 - 77\% probability. In
+contrast to the INCEPTION\citep{CPR2023a} original conclusion of similar
+survival effects between the two treatments, this re-analysis suggests
+only a modest probability of approximately 15\% that survival
+probabilities between the two techniques are within a reasonably
+acceptable equivalency range. Additional uncertainty about the benefits
+or harms of eCPR do arise with the per protocol analysis, both with and
+without informative priors.
 
-The INCEPTION researchers have addressed an important clinical question
-in the most challenging of research environments and are to be
-congratulated on their trial design, its execution, and a nuanced
-discussion. However the natural constraints of standard statistical
-analyses, also known as null hypothesis significance testing, limits the
-quantitative appreciation of their data, and prevents a full and
-comprehensive exploitation and updating of past knowledge. For trials
-such as INCEPTION that fail to meet statistical significance and are
-often incorrectly thought of as ``negative'' trials, null hypothesis
-significance testing favors confusion ``between absence of evidence and
-evidence of absence''\citep{RN6765}. Rather than dichotomizing results
-into statistical significance or not, with an obligatory loss of
-information and understanding, Bayesian analysis concentrates on
-estimation of key outcome differences with direct probability measures
-of their uncertainty.
+The INCEPTION\citep{CPR2023a} researchers have addressed an important
+clinical question in the most challenging of research environments and
+are to be congratulated on their trial design, its execution, and a
+nuanced discussion. However the natural constraints of standard
+statistical analyses, also known as null hypothesis significance
+testing, limits the quantitative appreciation of their data, and
+prevents a full and comprehensive exploitation and updating of past
+knowledge. For trials such as INCEPTION\citep{CPR2023a} that fail to
+meet statistical significance and are often incorrectly thought of as
+``negative'' trials, null hypothesis significance testing favors
+confusion ``between absence of evidence and evidence of
+absence''\citep{RN6765}. Rather than dichotomizing results into
+statistical significance or not, with an obligatory loss of information
+and understanding, Bayesian analysis concentrates on estimation of key
+outcome differences with direct probability measures of their
+uncertainty.
 
@@
 subgroups. Also for the per protocol analysis, this data was only
-available for survival and not the primary INCEPTION outcome that
-included survival with intact neurological status. The threshold choices
-for superiority and equivalence may be seen as arbitrary, although the
-analyses can be easily repeated for different choices. Given the
-existence of only 3 trials estimating the very divergent baseline cCPR
-rates could not be reliably modeled. We choose not to perform a
-traditional Bayesian meta-analysis with a semi-informative prior for the
-between-study variation because our goal was to illustrate how Bayesian
-principles could be informative when applied to the analysis and
-interpretation of new trial data, in situations both with and without
-previous knowledge.
+available for survival and not the primary INCEPTION\citep{CPR2023a}
+outcome that included survival with intact neurological status. The
+threshold choices for superiority and equivalence may be seen as
+arbitrary, although the analyses can be easily repeated for different
+choices. Given the existence of only 3 trials estimating the very
+divergent baseline cCPR rates could not be reliably modeled. We choose
+not to perform a traditional Bayesian meta-analysis with a
+semi-informative prior for the between-study variation because our goal
+was to illustrate how Bayesian principles could be informative when
+applied to the analysis and interpretation of new trial data, in
+situations both with and without previous knowledge.
 
